---
title: "Univariate and Multivariate Analysis of Dental Education Survey"
subtitle: "REVISED 2026-09 (EJDE major revision): minimum changes to the submitted script - all edits marked # REVISION"
params:
  boot_B: 200
author: "M Sc. Gregorio"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, knitr.kable.NA = "")
# ---- setup: repos + paquetes ----
# Define CRAN al knit (no interactivo)
if (is.null(getOption("repos")) || getOption("repos")["CRAN"] %in% c("@CRAN@", NA)) {
  options(repos = c(CRAN = "https://cran.rstudio.com/"))
}
```

```{r}
# ============================================================
# 0) Packages
# ============================================================
pkgs <- c(
  "readxl", "dplyr", "tidyr", "purrr", "stringr", "janitor",
  "knitr", "tibble"
)

to_install <- pkgs[!pkgs %in% rownames(installed.packages())]
if (length(to_install) > 0) install.packages(to_install)

invisible(lapply(pkgs, library, character.only = TRUE))
```


```{r}
# ============================================================
# REVISION (major revision, 2026-09) - DATA SOURCE
# The submitted analysis loaded the pre-built objects Data/survey_db.rds and Data/db_by_instrument.rds.
# Those objects contained (i) the responses of an excluded student record (#45) instead of the included student #59 and
# (ii) a spreadsheet copy-down error in the four educator adoption/maintenance items (all 16 educators = educator 1).
# Both objects are therefore rebuilt here directly from the original survey export (verbatim responses), keeping
# EXACTLY the same structure (6 sheets; identical column names, labels and segment maps) so that every downstream
# chunk of the original script runs unchanged. See docs/revision_note.md for the public revision record.
# ============================================================
set.seed(20260929)                                               # REVISION 2026: fixed seed (bootstrap section at the end)
# Repository root: this Rmd is stored in the analysis/ subdirectory
analysis_dir <- normalizePath(dirname(knitr::current_input(dir = TRUE)), winslash = "/", mustWork = FALSE)
proj_dir <- normalizePath(file.path(analysis_dir, ".."), winslash = "/", mustWork = FALSE)
data_dir <- file.path(proj_dir, "data")
restricted_out <- file.path(proj_dir, "restricted")
dir.create(restricted_out, showWarnings = FALSE, recursive = TRUE)
f_raw <- list.files(data_dir, pattern = "raw.?data.*\\.xlsx$", ignore.case = TRUE, full.names = TRUE, recursive = TRUE)
pub_items <- file.path(data_dir, "processed", "CCI_final_participant_level_deidentified.csv"); pub_adh <- file.path(data_dir, "processed", "CCI_final_adherence_deidentified.csv")
# REVISION 2026: two input modes. "raw" = the authors' raw workbook (verbatim responses; restricted, contains identifiers in other sheets);
# "public" = the de-identified FINAL files exported by this chunk in raw mode (76 participants, 60 selected D-OHRs; coded items, school code, role;
# no names, e-mails, institutions, free text, age or sex; sequential participant codes S01-S60 / E01-E16).
DATA_MODE <- if (length(f_raw) == 1) "raw" else if (file.exists(pub_items) && file.exists(pub_adh)) "public" else
  stop("REVISION check: provide either the raw workbook ('DB-Raw data ... CTR-26-OA-0061.xlsx') or data/processed/CCI_*_deidentified.csv")
if (DATA_MODE == "raw" && !"Raw data quest general-Info" %in% readxl::excel_sheets(f_raw)) stop("REVISION check: sheet 'Raw data quest general-Info' not found")
message("REVISION 2026: data mode = ", DATA_MODE)

item_map <- tibble::tribble(
  ~instrument, ~code, ~name, ~label, ~segment,
  "CCI", "CCI_B15a", "regarding_the_use_of_the_cci_system_for_caries_management_to_what_extent_did_you_consider_it_relevant_pertinent_for_the_project", "Regarding the use of the CCI system for caries management, to what extent did you consider it relevant/pertinent for the project?", "Relevance",
  "CCI", "CCI_B15b", "regarding_the_use_of_the_cci_system_for_caries_management_to_what_extent_would_you_consider_it_relevant_pertinent_for_the_dental_schools_clinics", "Regarding the use of the CCI system for caries management, to what extent would you consider it relevant/pertinent for the dental school's clinics?", "Relevance",
  "CCI", "CCI_B17", "regarding_the_use_of_the_cci_system_for_caries_management_to_what_extent_did_you_consider_there_was_compatibility_between_it_and_the_dental_schools_curricular_proposal", "Regarding the use of the CCI system for caries management, to what extent did you consider there was compatibility between it and the dental school's curricular proposal?", "Relevance",
  "CCI", "CCI_B18", "regarding_the_use_of_the_cci_system_for_caries_management_to_what_extent_would_you_consider_there_would_be_compatibility_between_it_and_the_dental_schools_curricular_proposal_for_the_clinics", "Regarding the use of the CCI system for caries management, to what extent would you consider  there would be compatibility between it and the dental school's curricular proposal for the clinics?", "Relevance",
  "CCI", "CCI_B19a", "regarding_the_use_of_the_cci_system_for_caries_management_to_what_extent_did_you_consider_it_useful_for_the_project", "Regarding the use of the CCI system for caries management, to what extent did you consider it useful for the project?", "Relevance",
  "CCI", "CCI_B19b", "regarding_the_use_of_the_cci_system_for_caries_management_to_what_extent_would_you_consider_it_useful_for_the_dental_schools_clinics", "Regarding the use of the CCI system for caries management, to what extent would you consider it useful for the dental school's clinics?", "Relevance",
  "CCI", "CCI_B1", "how_trained_did_you_feel_to_use_cci", "How trained did you feel to use CCI?", "Feasibility",
  "CCI", "CCI_B2", "to_properly_apply_cci_to_what_extent_did_you_consider_you_had_enough_time", "To properly apply CCI, to what extent did you consider you had enough time?", "Feasibility",
  "CCI", "CCI_B3a", "to_properly_apply_cci_to_what_extent_did_you_consider_you_had_the_required_radiographs", "To properly apply CCI, to what extent did you consider you had the required radiographs?", "Feasibility",
  "CCI", "CCI_B3b", "to_properly_apply_cci_to_what_extent_did_you_consider_you_had_the_required_dental_materials", "To properly apply CCI, to what extent did you consider you had the required dental materials?", "Feasibility",
  "CCI", "CCI_B4", "how_implementable_did_you_consider_cci_in_the_project", "How implementable did you consider CCI in the project?", "Feasibility",
  "CCI", "CCI_B5", "how_implementable_would_you_consider_cci_in_the_dental_schools_clinics", "How implementable would you consider CCI in the dental school's clinics?", "Feasibility",
  "CCI", "CCI_B6", "how_easy_did_you_find_applying_cci", "How easy did you find applying CCI?", "Acceptance",
  "CCI", "CCI_B7", "how_motivated_did_you_feel_applying_cci", "How motivated did you feel applying CCI?", "Acceptance",
  "CCI", "CCI_B8", "how_satisfied_did_you_feel_using_cci", "How satisfied did you feel using CCI?", "Acceptance",
  "CCI", "CCI_B21", "to_what_extent_did_you_adopt_cci", "To what extent did you adopt CCI?", "Adoption",
  "CCI", "CCI_B9", "to_what_degree_of_fidelity_did_you_apply_cci", "To what degree of fidelity did you apply CCI?", "Fidelity",
  "CCI", "CCI_B10", "to_what_degree_of_fidelity_would_you_consider_it_possible_to_apply_cci_in_the_dental_schools_clinics", "To what degree of fidelity would you consider it possible to apply CCI in the dental schools' clinics?", "Fidelity",
  "CCI", "CCI_B11a", "to_what_degree_of_fidelity_did_you_assess_patients_caries_risk_1d", "To what degree of fidelity did you assess patients<U+2019> caries risk (1D)?", "Fidelity",
  "CCI", "CCI_B11b", "to_what_degree_of_fidelity_would_you_consider_it_possible_to_assess_patients_caries_risk_1d_in_the_dental_school_s_clinics", "To what degree of fidelity would you consider it possible to assess patients<U+2019> caries risk (1D) in the dental school<U+2019>s clinics?", "Fidelity",
  "CCI", "CCI_B12a", "to_what_degree_of_fidelity_did_you_detected_and_assessed_caries_severity_and_activity_in_patients_2d", "To what degree of fidelity did you detected and assessed caries severity and activity in patients (2D)?", "Fidelity",
  "CCI", "CCI_B12b", "to_what_degree_of_fidelity_would_you_consider_it_possible_to_assess_caries_severity_and_activity_in_patients_2d_in_the_dental_school_s_clinics", "To what degree of fidelity would you consider it possible to assess caries severity and activity in patients (2D) in the dental school<U+2019>s clinics?", "Fidelity",
  "CCI", "CCI_B13a", "to_what_degree_of_fidelity_did_you_decided_a_personalised_care_plan_at_the_individual_and_tooth_surface_level_for_patients_3d", "To what degree of fidelity did you decided a personalised care plan at the individual and tooth surface level for patients (3D)?", "Fidelity",
  "CCI", "CCI_B13b", "to_what_degree_of_fidelity_would_you_consider_it_possible_to_decide_a_personalised_care_plan_at_the_individual_and_tooth_surface_level_for_patients_3d_in_the_dental_school_s_clinics", "To what degree of fidelity would you consider it possible to decide a personalised care plan at the individual and tooth surface level for patients (3D) in the dental school<U+2019>s clinics?", "Fidelity",
  "CCI", "CCI_B14a", "to_what_degree_of_fidelity_did_you_developed_a_preventive_and_tooth_preserving_management_of_patients_4d", "To what degree of fidelity did you developed a preventive and tooth-preserving management of patients (4D)?", "Fidelity",
  "CCI", "CCI_B14b", "to_what_degree_of_fidelity_would_you_consider_it_possible_to_develop_a_preventive_and_tooth_preserving_management_of_patients_4d_in_the_dental_school_s_clinics", "To what degree of fidelity would you consider it possible to develop a preventive and tooth-preserving management of patients (4D) in the dental school<U+2019>s clinics?", "Fidelity",
  "CCI", "CCI_B22", "to_what_extent_was_the_use_of_cci_maintained", "To what extent was the use of CCI maintained?", "Maintenance",
  "HxC", "DOHR_B16a", "regarding_the_use_of_the_digital_oral_health_record_to_what_extent_did_you_consider_it_relevant_pertinent_for_the_project", "Regarding the use of the digital oral health record, to what extent did you consider it relevant/pertinent for the project?", "Relevance",
  "HxC", "DOHR_B16b", "regarding_the_use_of_the_digital_oral_health_record_to_what_extent_would_you_consider_it_relevant_pertinent_for_the_dental_schools_clinics", "Regarding the use of the digital oral health record, to what extent would you consider it relevant/pertinent for the dental school's clinics?", "Relevance",
  "HxC", "DOHR_B17", "regarding_the_use_of_the_digital_oral_health_record_to_what_extent_did_you_consider_there_was_compatibility_between_it_and_the_dental_schools_curricular_proposal", "Regarding the use of the digital oral health record, to what extent did you consider there was compatibility between it and the dental school's curricular proposal?", "Relevance",
  "HxC", "DOHR_B18", "regarding_the_use_of_the_digital_oral_health_record_to_what_extent_would_you_consider_there_would_be_compatibility_between_it_and_the_dental_schools_curricular_proposal_for_the_clinics", "Regarding the use of the digital oral health record, to what extent would you consider there would be compatibility between it and the dental school's curricular proposal for the clinics?", "Relevance",
  "HxC", "DOHR_B20a", "regarding_the_use_of_the_digital_oral_health_record_to_what_extent_did_you_consider_it_useful_for_the_project", "Regarding the use of the digital oral health record, to what extent did you consider it useful for the project?", "Relevance",
  "HxC", "DOHR_B20b", "regarding_the_use_of_the_digital_oral_health_record_to_what_extent_would_you_consider_it_useful_for_the_dental_schools_clinics", "Regarding the use of the digital oral health record, to what extent would you consider it useful for the dental school's clinics?", "Relevance",
  "HxC", "DOHR_B1", "how_trained_did_you_feel_to_use_the_d_ohr", "How trained did you feel to use the D-OHR?", "Feasibility",
  "HxC", "DOHR_B2", "to_what_extent_did_you_consider_you_had_enough_time_to_properly_apply_the_d_ohr", "To what extent did you consider you had enough time to properly apply the D-OHR?", "Feasibility",
  "HxC", "DOHR_B4", "how_implementable_did_you_consider_the_d_ohr_for_the_project", "How implementable did you consider the D-OHR for the project?", "Feasibility",
  "HxC", "DOHR_B5", "how_implementable_would_you_consider_the_d_ohr_for_the_dental_school_s_clinics", "How implementable would you consider the D-OHR for the dental school<U+2019>s clinics?", "Feasibility",
  "HxC", "DOHR_B6", "how_easy_did_you_find_applying_the_d_ohr", "How easy did you find applying the D-OHR?", "Acceptance",
  "HxC", "DOHR_B7", "how_motivated_did_you_feel_applying_d_ohr", "How motivated did you feel applying D-OHR?", "Acceptance",
  "HxC", "DOHR_B8", "how_satisfied_did_you_feel_using_d_ohr", "How satisfied did you feel using D-OHR?", "Acceptance",
  "HxC", "DOHR_B21", "to_what_extent_did_you_adopt_the_d_ohr", "To what extent did you adopt the D-OHR?", "Adoption",
  "HxC", "DOHR_B9", "to_what_degree_of_fidelity_did_you_complete_in_the_d_ohr_the_cci_4d_cycle", "To what degree of fidelity did you complete in the D-OHR the CCI 4D-cycle?", "Fidelity",
  "HxC", "DOHR_B10", "to_what_degree_of_fidelity_would_you_consider_it_possible_to_use_the_d_ohr_for_caries_care_dental_school_s_clinics", "To what degree of fidelity would you consider it possible to use the D-OHR for caries care dental school<U+2019>s clinics?", "Fidelity",
  "HxC", "DOHR_B22", "to_what_extent_was_the_use_of_the_d_ohr_maintained", "To what extent was the use of the D-OHR maintained?", "Maintenance",
  "TEI", "TEI_B27", "how_much_do_you_consider_that_your_knowledge_regarding_the_current_diagnosis_and_management_of_caries_increased", "How much do you consider that your knowledge regarding the current diagnosis and management of caries increased?", "TEI",
  "TEI", "TEI_B28", "how_much_do_you_consider_that_cci_facilitates_a_more_structured_decision_making_in_the_management_of_caries", "How much do you consider that CCI facilitates a more structured decision-making in the management of caries?", "TEI",
  "TEI", "TEI_B29", "how_much_do_you_think_that_the_cci_approaches_helped_patients_become_more_aware_of_achieving_better_long_term_oral_health", "How much do you think that the CCI approaches helped patients become more aware of achieving better long-term oral health?", "TEI",
  "TEI", "TEI_B30", "how_much_do_you_think_that_the_cci_approaches_helped_raise_your_awareness_about_achieving_better_long_term_oral_health_outcomes_for_your_patients", "How much do you think that the CCI approaches helped raise your awareness about achieving better long-term oral health outcomes for your patients?", "TEI",
  "TEI", "TEI_B31", "how_easy_was_it_learning_to_use_the_digital_oral_health_record", "How easy was it learning to use the digital oral health record?", "TEI",
  "TEI", "TEI_B34", "to_what_extent_would_you_recommend_your_colleagues_using_cci_in_the_practice", "To what extent would you recommend your colleagues using CCI in the practice?", "TEI",
  "TEI", "TEI_B35", "to_what_extent_would_you_recommend_your_colleagues_using_the_digital_oral_health_record_in_the_practice", "To what extent would you recommend your colleagues using the digital oral health record in the practice?", "TEI"
)
item_map$segment[item_map$segment == "Acceptance"] <- "Usability"   # REVISION 2026: manuscript term (source label 'Aceptación'; Methods/Table 2 = 'Usability')

# questionnaire order of the item columns in the export (CCI and D-OHR interleaved) and verbatim-label recoding (1-5)
raw_codes <- c("CCI_B1","DOHR_B1","CCI_B2","DOHR_B2","CCI_B3a","CCI_B3b","CCI_B4","DOHR_B4","CCI_B5","DOHR_B5","CCI_B6","DOHR_B6","CCI_B7","DOHR_B7",
  "CCI_B8","DOHR_B8","CCI_B9","DOHR_B9","CCI_B10","DOHR_B10","CCI_B11a","CCI_B11b","CCI_B12a","CCI_B12b","CCI_B13a","CCI_B13b","CCI_B14a","CCI_B14b",
  "CCI_B15a","CCI_B15b","DOHR_B16a","DOHR_B16b","CCI_B17","DOHR_B17","CCI_B18","DOHR_B18","CCI_B19a","CCI_B19b","DOHR_B20a","DOHR_B20b","CCI_B21","DOHR_B21","CCI_B22","DOHR_B22",
  paste0("TEI_B", c(27:31, 34, 35)))
RAW_ITEM_COLS <- c(17:60, 65:69, 72:73)
recode_likert <- function(x) { x0 <- stringi::stri_trans_general(stringr::str_to_lower(stringr::str_squish(x)), "Latin-ASCII")
  dplyr::case_when(x0 %in% c("nada","nada facil","aumento nada","facilita nada","ayudaron nada") ~ 1,
                   x0 %in% c("leve","poco facil","aumento poco","facilita poco","ayudaron poco") ~ 2,
                   x0 %in% c("moderadamente","moderadamente facil","aumento moderadamente","facilita moderadamente","ayudaron moderadamente") ~ 3,
                   x0 %in% c("sustancialmente","bastante facil","aumento bastante","facilita bastante","ayudaron bastante") ~ 4,
                   x0 %in% c("completamente","totalmente","muy facil","aumento mucho","facilita mucho","ayudaron mucho") ~ 5, TRUE ~ NA_real_) }
if (DATA_MODE == "raw") {
raw0 <- suppressMessages(readxl::read_excel(f_raw, sheet = "Raw data quest general-Info", col_names = FALSE, col_types = "text", .name_repair = "minimal"))
hdr <- stringr::str_squish(unlist(raw0[1, ])); raw0 <- raw0[-1, ]; raw0 <- raw0[!is.na(raw0[[1]]), ]
if (hdr[17] != "el sistema CCI?" || !grepl("^B27", hdr[65]) || !grepl("^B35", hdr[73])) stop("REVISION check: unexpected column layout of the raw export")
lik <- as.data.frame(lapply(raw0[, RAW_ITEM_COLS], recode_likert)); names(lik) <- raw_codes
PARTICIPANTS <- tibble::tibble(uid = as.integer(raw0[[1]]), consent = raw0[[2]], ds = raw0[[3]], sex = raw0[[4]], age = as.numeric(raw0[[5]]),
                               respondent_group = dplyr::recode(raw0[[6]], Estudiante = "Students", Docente = "Educators")) %>%
  dplyr::mutate(n = as.character(ifelse(respondent_group == "Educators", uid - 72L, uid))) %>% dplyr::bind_cols(lik)
} else {
  PARTICIPANTS <- tibble::as_tibble(utils::read.csv(pub_items, stringsAsFactors = FALSE, check.names = FALSE)) %>%
    dplyr::transmute(participant_code, n = as.character(as.integer(sub("^[SE]", "", participant_code))), uid = NA_integer_, consent = "Si",
                     ds = dental_school, sex = NA_character_, age = NA_real_, respondent_group, dplyr::across(dplyr::all_of(raw_codes)))
}
# ---- REVISION: structural validation (stop on any violation) ----
if (DATA_MODE == "raw") stopifnot(nrow(PARTICIPANTS) == 77, sum(PARTICIPANTS$respondent_group == "Students") == 61, sum(PARTICIPANTS$respondent_group == "Educators") == 16)
stopifnot(!anyDuplicated(paste(PARTICIPANTS$respondent_group, PARTICIPANTS$n)),
          all(PARTICIPANTS$consent == "Si"), all(PARTICIPANTS$ds %in% paste0("DS", 1:7)), setequal(unique(PARTICIPANTS$ds), paste0("DS", 1:7)),
          !anyNA(PARTICIPANTS[, raw_codes]), all(as.matrix(PARTICIPANTS[, raw_codes]) %in% 1:5),
          nrow(item_map) == 51, setequal(item_map$code, raw_codes), sum(item_map$instrument == "CCI") == 27, sum(item_map$instrument == "HxC") == 17)

# ---- REVISION 2026: validated final sample ----
# Author-confirmed student flow: 71 initial students -> 7 withdrew + 4 had no patient under their care -> 60 completed.
# The verbatim sheet still contains student #51, one of the four students without patients (author-confirmed; 0 patients in the
# adherence source sheet), who answered the questionnaire. #51 is therefore removed here; it is not part of the final analytic sample.
read_adherence <- function() {   # one row per student in the adherence audit sheet (one D-OHR per student)
  if (DATA_MODE == "public") return(tibble::as_tibble(utils::read.csv(pub_adh, stringsAsFactors = FALSE)) %>%
    dplyr::transmute(ds = dental_school, n_patients, student = as.integer(sub("^S", "", participant_code)), not_completed = !dohr_assessable,
                     recall, fluoride, noc_active, monitor_inactive))
  ad0 <- suppressMessages(readxl::read_excel(f_raw, sheet = "FidelitAssess-patients Info", col_names = FALSE, col_types = "text", .name_repair = "minimal"))[-(1:2), ]
  ad0 <- ad0[!is.na(ad0[[1]]) & grepl("^DS[0-9]$", ad0[[1]]), ]
  to01 <- function(z) { z <- tolower(trimws(z)); ifelse(z %in% c("0", "1"), as.numeric(z), NA_real_) }     # 'na', 'Na', '?' = not applicable
  tibble::tibble(ds = ad0[[1]], n_patients = as.numeric(ad0[[2]]), student = as.integer(ad0[[5]]), not_completed = !is.na(ad0[[7]]) & tolower(trimws(ad0[[7]])) == "x",
                 recall = to01(ad0[[8]]), fluoride = to01(ad0[[9]]), noc_active = to01(ad0[[10]]), monitor_inactive = to01(ad0[[11]])) }
ADH_SRC <- read_adherence()
patients_per_student <- stats::setNames(ADH_SRC$n_patients, as.character(ADH_SRC$student))
PARTICIPANTS_RAW <- PARTICIPANTS
stopifnot(setequal(names(patients_per_student), PARTICIPANTS_RAW$n[PARTICIPANTS_RAW$respondent_group == "Students"]), !anyDuplicated(names(patients_per_student)))
NO_PATIENT_IDS <- names(patients_per_student)[patients_per_student == 0]
PARTICIPANTS <- PARTICIPANTS %>% dplyr::filter(!(respondent_group == "Students" & n %in% NO_PATIENT_IDS))
if (DATA_MODE == "raw") stopifnot(identical(NO_PATIENT_IDS, "51")) else stopifnot(length(NO_PATIENT_IDS) == 0)
stopifnot(nrow(PARTICIPANTS) == 76, sum(PARTICIPANTS$respondent_group == "Students") == 60,
          sum(PARTICIPANTS$respondent_group == "Educators") == 16, dplyr::n_distinct(paste(PARTICIPANTS$respondent_group, PARTICIPANTS$n)) == 76,
          all(table(PARTICIPANTS$ds[PARTICIPANTS$respondent_group == "Students"]) == c(10, 8, 5, 7, 9, 8, 13)))
# ---- REVISION 2026: participant codes and de-identified FINAL public files ----
if (DATA_MODE == "raw") {
  PARTICIPANTS <- PARTICIPANTS %>% dplyr::arrange(dplyr::desc(respondent_group), as.integer(n)) %>% dplyr::group_by(respondent_group) %>%
    dplyr::mutate(participant_code = sprintf("%s%02d", ifelse(respondent_group == "Students", "S", "E"), dplyr::row_number())) %>% dplyr::ungroup()
  dir.create(file.path(data_dir, "processed"), showWarnings = FALSE, recursive = TRUE)
  utils::write.csv(PARTICIPANTS %>% dplyr::transmute(participant_code, respondent_group, dental_school = ds, dplyr::across(dplyr::all_of(raw_codes))),
                   file.path(data_dir, "processed", "CCI_final_participant_level_deidentified.csv")
  code_of_student <- stats::setNames(PARTICIPANTS$participant_code[PARTICIPANTS$respondent_group == "Students"], PARTICIPANTS$n[PARTICIPANTS$respondent_group == "Students"])
  utils::write.csv(ADH_SRC %>% dplyr::filter(n_patients > 0) %>% dplyr::transmute(participant_code = unname(code_of_student[as.character(student)]), dental_school = ds, n_patients,
                   dohr_assessable = !not_completed, recall, fluoride, noc_active, monitor_inactive) %>% dplyr::arrange(participant_code),
                   file.path(data_dir, "processed", "CCI_final_adherence_deidentified.csv")
  utils::write.csv(PARTICIPANTS %>% dplyr::transmute(participant_code, study_record_id = n, respondent_group), file.path(restricted_out, "participant_code_to_study_id_RESTRICTED.csv"), row.names = FALSE)
  stopifnot(!anyNA(code_of_student))
}

# ---- REVISION: rebuild survey_db with the original structure ----
make_sheet <- function(grp_label, grp_value, ins) {
  im <- item_map[item_map$instrument == ins, ]; p <- PARTICIPANTS[PARTICIPANTS$respondent_group == grp_value, ]
  dat <- data.frame(n = p$n, ds = p$ds, stringsAsFactors = FALSE)
  for (i in seq_len(nrow(im))) dat[[im$name[i]]] <- as.character(p[[im$code[i]]])
  dat$respondent_group <- grp_value; dat$instrument <- ins
  list(sheet = paste(grp_label, ins), data = dat,
       seg_map = c(n = "Unsegmented", ds = "Unsegmented", stats::setNames(im$segment, im$name)),
       label_map = c(n = "n", ds = "DS", stats::setNames(im$label, im$name)),
       meta = list(respondent_group = grp_value, instrument = ins))
}
survey_db <- list(`Students CCI` = make_sheet("Students", "Students", "CCI"), `Students HxC` = make_sheet("Students", "Students", "HxC"),
                  `Students TEI` = make_sheet("Students", "Students", "TEI"), `Teachers CCI` = make_sheet("Teachers", "Educators", "CCI"),
                  `Teachers HxC` = make_sheet("Teachers", "Educators", "HxC"), `Teachers TEI` = make_sheet("Teachers", "Educators", "TEI"))
db_by_instrument <- list(CCI = dplyr::bind_rows(survey_db[["Students CCI"]]$data, survey_db[["Teachers CCI"]]$data),
                         HxC = dplyr::bind_rows(survey_db[["Students HxC"]]$data, survey_db[["Teachers HxC"]]$data),
                         TEI = dplyr::bind_rows(survey_db[["Students TEI"]]$data, survey_db[["Teachers TEI"]]$data))   # per-instrument (different people): row-binding groups is correct
str(survey_db, max.level = 1)

# ---- REVISION: FINAL ANALYTIC DATABASE (one row per participant) + machine-readable dictionary ----
FINAL_DB <- PARTICIPANTS %>% dplyr::transmute(participant_id = participant_code, study_record_id = as.integer(n),
                                              raw_record_id = uid, respondent_group, dental_school = ds, sex, age) %>%
  dplyr::bind_cols(PARTICIPANTS[, item_map$code])
DICTIONARY <- item_map %>% dplyr::transmute(variable = code, instrument = ifelse(instrument == "HxC", "D-OHR", instrument), domain = segment, question = label,
                                            legacy_column_name = name, coding = "1-5 ordinal (1 = none/nothing ... 5 = complete/very); recoded from verbatim Spanish labels",
                                            reverse_coded = "no") %>%
  dplyr::bind_rows(tibble::tibble(variable = c("participant_id", "study_record_id", "raw_record_id", "respondent_group", "dental_school", "sex", "age"), instrument = "-", domain = "identifier/demographic",
                                  question = c("Participant code (S01-S60 students, E01-E16 educators; sequential)", "Study record identifier (an identifier, not a count; restricted)", "Record identifier in the raw export (restricted)", "Students / Educators", "DS1-DS7 (manuscript coding; institution identity confidential)", "Femenino / Masculino", "Years"),
                                  legacy_column_name = c("-", "N° (role-specific)", "N°", "respondent_group", "ds", "A3", "A4"), coding = "-", reverse_coded = "-"))
stopifnot(nrow(FINAL_DB) == 76, !anyDuplicated(FINAL_DB$participant_id), ncol(FINAL_DB) == 7 + 51)
saveRDS(FINAL_DB, file.path(restricted_out, "FINAL_ANALYTIC_DATABASE.rds"))
writexl::write_xlsx(
  list(
    FINAL_ANALYTIC_DATABASE = as.data.frame(FINAL_DB),
    DATA_DICTIONARY = as.data.frame(DICTIONARY)
  ),
  file.path(restricted_out, "FINAL_ANALYTIC_DATABASE.xlsx")
)
write.csv(DICTIONARY, file.path(restricted_out, "DATA_DICTIONARY.csv"), row.names = FALSE)
```

<!-- REVISION: original data-loading chunk (replaced by the chunk above), kept for traceability:
# # ============================================================
# # Cargar la base de datos desde el archivo RDS
# # ============================================================
# # Ruta donde se guarda el archivo RDS
# data_path <- "Data/survey_db.rds"
# 
# # Cargar el archivo RDS
# survey_db <- readRDS(file = data_path)
# 
# # Verificar si se cargó correctamente
# str(survey_db)  # Esto mostrará la estructura del objeto cargado
# 
# # ============================================================
# # Cargar la base de datos desde el archivo RDS
# # ============================================================
# # Ruta donde se guarda el archivo RDS
# data_path <- "Data/db_by_instrument.rds"
# 
# # Cargar el archivo RDS
# db_by_instrument <- readRDS(file = data_path)
# 
# # Verificar si se cargó correctamente
# str(db_by_instrument)  # Esto mostrará la estructura del objeto cargado
-->

```{r}
# ============================================================
# SECTION 0) Helpers (Mode & IQR + safe item selection)
# ============================================================
mode_numeric <- function(x) {
  x <- suppressWarnings(as.numeric(as.character(x)))
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_real_)
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

iqr_numeric <- function(x) {
  x <- suppressWarnings(as.numeric(as.character(x)))
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_real_)
  IQR(x)
}

# Given a dataframe and a segment map (named by cleaned column names),
# return ONLY item columns to analyze, excluding:
# - col1 (id) and col2 (DS) ALWAYS
# - respondent_group and instrument ALWAYS
# - any explicitly excluded columns you pass (optional)
get_item_columns <- function(df, seg_map, extra_exclude = character()) {
  stopifnot(ncol(df) >= 2)

  col1 <- names(df)[1]
  col2 <- names(df)[2]

  exclude <- unique(c(col1, col2, "respondent_group", "instrument", extra_exclude))
  exclude <- exclude[exclude %in% names(df)]

  item_cols <- setdiff(names(df), exclude)

  # Keep only those that have a meaningful segment label
  seg_item <- seg_map[item_cols]
  keep <- !is.na(seg_item) &
    stringr::str_trim(seg_item) != "" &
    stringr::str_trim(seg_item) != "Unsegmented"

  item_cols[keep]
}

# ============================================================
# SECTION 1) Build ONE long dataset (from survey_db) with:
#   ds, respondent_group, instrument, segment, item, item_label, value
# Notes:
# - ds is taken from column 2 (always DS)
# - excludes col1=id, col2=DS from item list automatically
# - excludes respondent_group + instrument from item list automatically
# ============================================================
build_long_all <- function(survey_db) {

  one_sheet_to_long <- function(sheet_obj) {
    df        <- sheet_obj$data
    seg_map   <- sheet_obj$seg_map
    label_map <- sheet_obj$label_map

    stopifnot(ncol(df) >= 2)

    id_col <- names(df)[1]
    ds_col <- names(df)[2]

    item_cols <- get_item_columns(df, seg_map)

    if (length(item_cols) == 0) {
      return(tibble::tibble(
        ds = character(),
        respondent_group = character(),
        instrument = character(),
        segment = character(),
        item = character(),
        item_label = character(),
        value = numeric()
      ))
    }

    seg_item   <- seg_map[item_cols]
    label_item <- label_map[item_cols]

    df %>%
      dplyr::mutate(
        id = .data[[id_col]],
        ds = .data[[ds_col]]
      ) %>%
      dplyr::select(dplyr::all_of(c("ds", "respondent_group", "instrument", item_cols))) %>%
      tidyr::pivot_longer(
        cols      = dplyr::all_of(item_cols),
        names_to  = "item",
        values_to = "value"
      ) %>%
      dplyr::mutate(
        segment    = unname(seg_item[item]),
        item_label = unname(label_item[item]),
        value      = suppressWarnings(as.numeric(as.character(value)))
      ) %>%
      dplyr::filter(!is.na(segment))
  }

  purrr::map_dfr(survey_db, one_sheet_to_long) %>%
    dplyr::filter(
      !is.na(ds), ds != "",
      respondent_group %in% c("Students", "Educators"),
      instrument %in% c("CCI", "HxC", "TEI")
    )
}

# Build it (uses your existing survey_db from earlier)
all_long <- build_long_all(survey_db)

# Optional quick sanity checks
dplyr::count(all_long, instrument, respondent_group)
dplyr::count(all_long, ds, instrument)

# ============================================================
# SECTION 2) MODE & IQR by DS x Respondent_group x Instrument x Segment
# (This is the "by segment" summary, not item-level)
# ============================================================
mode_iqr_by_ds_resp_instr_segment <- all_long %>%
  dplyr::group_by(ds, respondent_group, instrument, segment, .drop = TRUE) %>%
  dplyr::summarise(
    n_nonmiss = sum(!is.na(value)),
    mode      = mode_numeric(value),
    iqr       = iqr_numeric(value),
    .groups   = "drop"
  ) %>%
  dplyr::arrange(instrument, ds, respondent_group, segment)

knitr::kable(
  mode_iqr_by_ds_resp_instr_segment,
  caption = "Mode & IQR by DS × Respondent group × Instrument, grouped by Segment",
  digits  = 2
)

# ============================================================
# SECTION 3) MODE & IQR by DS x Respondent_group x Instrument x Segment x Item
# (This is the item/column-level summary you requested)
# ============================================================
mode_iqr_by_ds_resp_instr_segment_item <- all_long %>%
  dplyr::group_by(ds, respondent_group, instrument, segment, item, item_label, .drop = TRUE) %>%
  dplyr::summarise(
    n_nonmiss = sum(!is.na(value)),
    mode      = mode_numeric(value),
    iqr       = iqr_numeric(value),
    .groups   = "drop"
  ) %>%
  dplyr::arrange(instrument, ds, respondent_group, segment, item_label)

knitr::kable(
  mode_iqr_by_ds_resp_instr_segment_item,
  caption = "Mode & IQR by DS × Respondent group × Instrument × Segment × Item",
  digits  = 2
)

# ============================================================
# SECTION 4) "All combinations" (optional): a small function
# that lets you compute Mode & IQR for ANY grouping subset.
#
# Examples:
#   - by instrument + segment
#   - by ds + instrument + item
#   - by respondent_group + segment + item
# ============================================================
mode_iqr_any_grouping <- function(long_df,
                                 group_vars = c("instrument", "segment"),
                                 include_item = FALSE) {

  needed <- c("ds","respondent_group","instrument","segment","item","item_label","value")
  stopifnot(all(group_vars %in% needed))
  stopifnot(all(needed %in% names(long_df)))

  if (include_item) {
    group_vars <- unique(c(group_vars, "item", "item_label"))
  }

  long_df %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars)), .drop = TRUE) %>%
    dplyr::summarise(
      n_nonmiss = sum(!is.na(value)),
      mode      = mode_numeric(value),
      iqr       = iqr_numeric(value),
      .groups   = "drop"
    ) %>%
    dplyr::arrange(dplyr::across(dplyr::all_of(group_vars)))
}

# ---- EXAMPLES of "all combinations" you might want ----

# A) By instrument x segment (collapsed across DS and respondent group)
ex_instr_segment <- mode_iqr_any_grouping(all_long, c("instrument","segment"), include_item = FALSE)

# B) By instrument x segment x item
ex_instr_segment_item <- mode_iqr_any_grouping(all_long, c("instrument","segment"), include_item = TRUE)

# C) By DS x instrument x segment
ex_ds_instr_segment <- mode_iqr_any_grouping(all_long, c("ds","instrument","segment"), include_item = FALSE)

# D) By DS x instrument x segment x item
ex_ds_instr_segment_item <- mode_iqr_any_grouping(all_long, c("ds","instrument","segment"), include_item = TRUE)

# E) By respondent_group x instrument x segment
ex_resp_instr_segment <- mode_iqr_any_grouping(all_long, c("respondent_group","instrument","segment"), include_item = FALSE)

# Print any example
knitr::kable(ex_resp_instr_segment, caption = "Example: Mode & IQR by Respondent × Instrument x Segment", digits = 2)

# Print any example
knitr::kable(ex_instr_segment, caption = "Example: Mode & IQR by Instrument × Segment", digits = 2)
```

```{r}
# ============================================================
# REAL Cronbach's alpha from survey_db (wide respondent x item)
# ============================================================

pkgs <- c("dplyr","purrr","stringr","tibble","psych")
to_install <- pkgs[!pkgs %in% rownames(installed.packages())]
if(length(to_install)) install.packages(to_install)
invisible(lapply(pkgs, library, character.only = TRUE))

# ------------------------------------------------------------
# Helper: safe psych alpha on a respondent x item matrix
# ------------------------------------------------------------
cronbach_psych_safe <- function(mat, n_min = 10) {
  mat <- as.data.frame(mat)

  # numeric coercion
  mat <- mat %>% mutate(across(everything(), ~ suppressWarnings(as.numeric(as.character(.x)))))

  # drop all-NA cols
  mat <- mat[, colSums(!is.na(mat)) > 0, drop = FALSE]

  k0 <- ncol(mat)
  if (k0 < 2) {
    return(tibble(
      n_complete = sum(complete.cases(mat)),
      k_items = k0,
      dropped_zero_var = 0L,
      alpha = NA_real_,
      note = "Alpha undefined: <2 items in segment."
    ))
  }

  # listwise complete cases (standard for alpha)
  mat_cc <- mat[complete.cases(mat), , drop = FALSE]
  n_cc <- nrow(mat_cc)

  if (n_cc < n_min) {
    return(tibble(
      n_complete = n_cc,
      k_items = k0,
      dropped_zero_var = 0L,
      alpha = NA_real_,
      note = paste0("Too few complete cases (n<", n_min, ").")
    ))
  }

  # drop zero-variance items (alpha becomes unstable / fails)
  v <- vapply(mat_cc, var, numeric(1), na.rm = TRUE)
  keep <- names(v)[is.finite(v) & v > 0]
  dropped <- k0 - length(keep)
  mat_cc <- mat_cc[, keep, drop = FALSE]

  if (ncol(mat_cc) < 2) {
    return(tibble(
      n_complete = n_cc,
      k_items = ncol(mat_cc),
      dropped_zero_var = dropped,
      alpha = NA_real_,
      note = "After removing zero-variance items, <2 items remain."
    ))
  }

  a <- tryCatch(
    psych::alpha(mat_cc, check.keys = FALSE, warnings = FALSE),
    error = function(e) NULL
  )

  if (is.null(a)) {
    return(tibble(
      n_complete = n_cc,
      k_items = ncol(mat_cc),
      dropped_zero_var = dropped,
      alpha = NA_real_,
      note = "psych::alpha failed."
    ))
  }

  tibble(
    n_complete = a$n.obs,
    k_items = a$nvar,
    dropped_zero_var = dropped,
    alpha = unname(a$total$raw_alpha),
    note = "OK"
  )
}

# ------------------------------------------------------------
# Main: compute alpha by SEGMENT, and optionally by DS
# Uses:
#  - respondent id = col 1 (as in your sheets)
#  - ds            = col 2
#  - items         = all other columns except respondent_group/instrument
#  - segment labels come from sheet_obj$seg_map
# ------------------------------------------------------------
alpha_by_segment_from_survey_db <- function(survey_db,
                                           by_ds = FALSE,
                                           n_min = 10) {

  one_sheet <- function(sheet_obj) {
    df <- sheet_obj$data
    seg_map <- sheet_obj$seg_map
    meta <- sheet_obj$meta

    # identify columns
    id_col <- names(df)[1]
    ds_col <- names(df)[2]

    # item columns are those in seg_map excluding id/ds/provenance
    exclude <- c(id_col, ds_col, "respondent_group", "instrument")
    item_cols <- setdiff(names(df), exclude)

    # Keep only segmented items (not Unsegmented)
    seg_for_item <- seg_map[item_cols]
    keep_items <- !is.na(seg_for_item) & stringr::str_trim(seg_for_item) != "" &
      stringr::str_trim(seg_for_item) != "Unsegmented"
    item_cols <- item_cols[keep_items]
    seg_for_item <- seg_for_item[keep_items]

    if (length(item_cols) == 0) return(tibble())

    # Build a long table of (item -> segment)
    item_seg_tbl <- tibble(item = item_cols, segment = unname(seg_for_item))

    # Base respondent table
    base <- df %>%
      transmute(
        respondent_id = .data[[id_col]],
        ds = .data[[ds_col]],
        respondent_group = .data[["respondent_group"]],
        instrument = .data[["instrument"]]
      )

    # Merge items back (wide stays wide, preserving respondent_id)
    wide_items <- df %>%
      select(all_of(c(id_col, ds_col, item_cols))) %>%
      rename(respondent_id = all_of(id_col),
             ds = all_of(ds_col))

    # now compute alpha per segment
    segments <- unique(item_seg_tbl$segment)

    purrr::map_dfr(segments, function(seg) {
      seg_items <- item_seg_tbl %>% filter(segment == seg) %>% pull(item)

      if (!by_ds) {
        mat <- wide_items %>% select(all_of(seg_items))
        out <- cronbach_psych_safe(mat, n_min = n_min)
        out %>% mutate(
          respondent_group = unique(base$respondent_group),
          instrument = unique(base$instrument),
          segment = seg,
          ds = NA_character_
        )
      } else {
        # by DS: split wide_items by ds and compute alpha in each DS
        wide_items %>%
          group_by(ds) %>%
          group_modify(~{
            mat <- .x %>% select(all_of(seg_items))
            cronbach_psych_safe(mat, n_min = n_min)
          }) %>%
          ungroup() %>%
          mutate(
            respondent_group = unique(base$respondent_group),
            instrument = unique(base$instrument),
            segment = seg
          )
      }
    })
  }

  purrr::map_dfr(survey_db, one_sheet) %>%
    # clean and order
    select(respondent_group, instrument, ds, segment, everything()) %>%
    arrange(respondent_group, instrument, ds, segment)
}

# ============================================================
# RUN IT
# ============================================================

# 1) Alpha by segment within each sheet group (Students/Educators × CCI/HxC/TEI)
alpha_segment <- alpha_by_segment_from_survey_db(survey_db, by_ds = FALSE, n_min = 10)

# 2) Alpha by segment within DS (this can be unstable if DS has small n)
alpha_segment_ds <- alpha_by_segment_from_survey_db(survey_db, by_ds = TRUE, n_min = 10)

knitr::kable(alpha_segment, digits = 3,
             caption = "Cronbach's alpha (psych) by Segment within Respondent group × Instrument")

knitr::kable(alpha_segment_ds, digits = 3,
             caption = "Cronbach's alpha (psych) by Segment within Respondent group × Instrument × DS")

```

```{r rev-zero-variance}
# REVISION 2026: zero-variance screen within each respondent group (the submitted text reported 4 invariant educator items;
# they were an artefact of the copy-down error in the coded spreadsheet)
zv <- purrr::imap_dfr(survey_db, function(sh, nm) { it <- get_item_columns(sh$data, sh$seg_map)
  tibble::tibble(sheet = nm, item = it, variance = vapply(sh$data[it], function(z) var(as.numeric(z)), numeric(1))) })
n_zero_var <- sum(zv$variance == 0)
knitr::kable(zv %>% dplyr::filter(variance == 0))
```

*REVISION 2026:* Prior to Multiple Factor Analysis, items were screened for zero variance within each respondent group. With the verified participant-level data, `r n_zero_var` items were invariant; therefore no item was excluded from the MFA. (Submitted text, withdrawn: "Prior to Multiple Factor Analysis, items presenting zero variance within specific respondent groups were identified and excluded, as they do not contribute to multivariate discrimination. Specifically, two items related to adoption and maintenance of CCI and two items related to digital oral health records (D-OHR) showed invariant responses among Educators, indicating unanimous agreement. These items were retained for descriptive analyses but excluded from MFA to ensure methodological validity.")

```{r}
pkgs <- c("dplyr","purrr","stringr","tibble","tidyr","FactoMineR","factoextra","missMDA","janitor")
to_install <- pkgs[!pkgs %in% rownames(installed.packages())]
if(length(to_install)) install.packages(to_install)
invisible(lapply(pkgs, library, character.only = TRUE))

```

```{r}
# --- reuso: columnas "items" excluyendo id, DS, respondent_group, instrument ---
get_item_columns <- function(df, seg_map, extra_exclude = character()) {
  stopifnot(ncol(df) >= 2)
  col1 <- names(df)[1]
  col2 <- names(df)[2]
  exclude <- unique(c(col1, col2, "respondent_group", "instrument", extra_exclude))
  exclude <- exclude[exclude %in% names(df)]
  item_cols <- setdiff(names(df), exclude)

  seg_item <- seg_map[item_cols]
  keep <- !is.na(seg_item) &
    stringr::str_trim(seg_item) != "" &
    stringr::str_trim(seg_item) != "Unsegmented"

  item_cols[keep]
}

# --- construir dataset ANCHO (wide) + grupos (segmentos) para MFA ---
build_mfa_input_from_sheetobj <- function(sheet_obj,
                                         keep_ds = FALSE,
                                         drop_constant = TRUE) {
  df        <- sheet_obj$data
  seg_map   <- sheet_obj$seg_map
  label_map <- sheet_obj$label_map

  id_col <- names(df)[1]
  ds_col <- names(df)[2]

  item_cols <- get_item_columns(df, seg_map)
  if (length(item_cols) < 2) return(NULL)

  # X: solo items (numéricos)
  X <- df %>%
    dplyr::select(dplyr::all_of(item_cols)) %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), ~ suppressWarnings(as.numeric(as.character(.x))))) %>%
    as.data.frame()

  # opcional: quitar columnas constantes (var=0) que fastidian MFA
  if (drop_constant) {
    v <- vapply(X, function(z) var(z, na.rm = TRUE), numeric(1))
    keep <- names(v)[is.finite(v) & v > 0]
    X <- X[, keep, drop = FALSE]
    item_cols <- intersect(item_cols, keep)
    if (length(item_cols) < 2) return(NULL)
  }

  # grupos = segmentos, en el orden de columnas actuales
  seg_vec <- unname(seg_map[item_cols])
  seg_vec[is.na(seg_vec) | str_trim(seg_vec)=="" ] <- "Unsegmented"

  # mantener solo segmentados
  keep2 <- str_trim(seg_vec) != "Unsegmented"
  X <- X[, keep2, drop = FALSE]
  seg_vec <- seg_vec[keep2]
  item_cols <- item_cols[keep2]
  if (ncol(X) < 2) return(NULL)

  # tamaños de grupo para MFA
  groups <- tibble(item = item_cols, segment = seg_vec) %>%
    count(segment, name = "ng") %>%
    arrange(segment)

  # reordenar columnas por segmento (mejor lectura)
  ord <- order(seg_vec, item_cols)
  X <- X[, ord, drop = FALSE]
  seg_vec <- seg_vec[ord]
  item_cols <- item_cols[ord]

  group_sizes <- as.integer(tibble(segment=seg_vec) %>% count(segment) %>% pull(n))
  group_names <- as.character(tibble(segment=seg_vec) %>% count(segment) %>% pull(segment))

  # metadatos fila (id/ds) por si quieres colorear después
  meta_rows <- df %>%
    transmute(
      respondent_id = .data[[id_col]],
      ds = .data[[ds_col]],
      respondent_group = .data[["respondent_group"]],
      instrument = .data[["instrument"]]
    )

  if (!keep_ds) meta_rows <- meta_rows %>% select(-ds)

  # labels bonitos (opcional)
  item_labels <- unname(label_map[item_cols])

  list(
    X = X,
    meta = meta_rows,
    group_sizes = group_sizes,
    group_names = group_names,
    item_cols = item_cols,
    item_labels = item_labels
  )
}

```

```{r}
# ============================================================
# MFA inputs por combo (Students/Educators × Instrument) ROBUSTO
# - Alinea columnas
# - Recalcula segmentación por columna buscando en TODAS las hojas del combo
# - Elimina columnas all-NA / var=0 / Unsegmented
# - Recalcula group_sizes SIEMPRE desde las columnas finales
# ============================================================

mfa_inputs_by_combo <- function(survey_db,
                               keep_ds = TRUE,
                               drop_constant = TRUE,
                               min_rows = 3,
                               min_cols = 2) {

  # ---- helper: saca X numérico de una hoja (sin ordenar por segmentos aquí) ----
  one_sheet_input <- function(sheet_obj, sheet_name) {
    df      <- sheet_obj$data
    seg_map <- sheet_obj$seg_map

    item_cols <- get_item_columns(df, seg_map)
    if (length(item_cols) < 2) return(NULL)

    X <- df %>%
      dplyr::select(dplyr::all_of(item_cols)) %>%
      dplyr::mutate(dplyr::across(dplyr::everything(),
                                  ~ suppressWarnings(as.numeric(as.character(.x))))) %>%
      as.data.frame()

    meta <- df %>%
      dplyr::transmute(
        respondent_id = .data[[names(df)[1]]],
        ds = .data[[names(df)[2]]],
        respondent_group = .data[["respondent_group"]],
        instrument = .data[["instrument"]]
      )

    if (!keep_ds) meta <- meta %>% dplyr::select(-ds)

    list(sheet = sheet_name, X = X, meta = meta)
  }

  # ---- 1) inputs por hoja ----
  inputs <- purrr::imap(survey_db, ~ one_sheet_input(.x, .y))
  inputs <- inputs[!vapply(inputs, is.null, logical(1))]

  if (length(inputs) == 0) return(list())

  # ---- 2) combos presentes ----
  combos <- purrr::map_dfr(inputs, ~ dplyr::distinct(.x$meta, respondent_group, instrument)) %>%
    dplyr::distinct()

  # ---- helper: segment para cada columna buscando en todas las hojas del combo ----
  get_seg_for_items <- function(items, sel_sheets, survey_db) {
    seg <- rep(NA_character_, length(items))
    names(seg) <- items

    for (sh in sel_sheets) {
      sm <- survey_db[[sh]]$seg_map
      hit <- sm[items]
      # rellenar donde aún es NA y aquí hay valor
      to_fill <- is.na(seg) & !is.na(hit)
      seg[to_fill] <- unname(hit[to_fill])
      if (!any(is.na(seg))) break
    }

    seg[is.na(seg) | stringr::str_trim(seg) == ""] <- "Unsegmented"
    seg
  }

  # ---- 3) construir dataset por combo ----
  out <- purrr::pmap(combos, function(respondent_group, instrument) {

    sel <- purrr::keep(inputs, ~{
      unique(.x$meta$respondent_group) == respondent_group &&
        unique(.x$meta$instrument) == instrument
    })

    if (length(sel) == 0) return(NULL)

    sel_sheets <- vapply(sel, `[[`, character(1), "sheet")

    # union de columnas
    all_items <- Reduce(union, lapply(sel, \(z) colnames(z$X)))

    # bind_rows alineando columnas
    X_all <- purrr::map_dfr(sel, \(z){
      X <- z$X
      miss <- setdiff(all_items, colnames(X))
      if (length(miss)) X[miss] <- NA_real_
      X <- X[, all_items, drop = FALSE]
      X
    })

    meta_all <- purrr::map_dfr(sel, \(z) z$meta)

    # --- limpieza columnas ---
    # (1) quitar columnas all-NA
    keep_not_allna <- colSums(!is.na(X_all)) > 0
    X_all <- X_all[, keep_not_allna, drop = FALSE]
    all_items2 <- colnames(X_all)

    if (ncol(X_all) < min_cols || nrow(X_all) < min_rows) return(NULL)

    # (2) segmentación por columna (buscando en todas las hojas del combo)
    seg_vec <- get_seg_for_items(all_items2, sel_sheets, survey_db)

    # (3) quitar Unsegmented
    keep_seg <- stringr::str_trim(seg_vec) != "Unsegmented"
    X_all <- X_all[, keep_seg, drop = FALSE]
    seg_vec <- seg_vec[keep_seg]

    if (ncol(X_all) < min_cols) return(NULL)

    # (4) quitar columnas constantes (var=0) si se pide
    if (drop_constant) {
      v <- vapply(X_all, function(z) var(z, na.rm = TRUE), numeric(1))
      keep_var <- is.finite(v) & v > 0
      X_all <- X_all[, keep_var, drop = FALSE]
      seg_vec <- seg_vec[keep_var]
      if (ncol(X_all) < min_cols) return(NULL)
    }

    # (5) ordenar columnas por segmento (para plots más claros)
    ord <- order(seg_vec, colnames(X_all))
    X_all <- X_all[, ord, drop = FALSE]
    seg_vec <- seg_vec[ord]

    # (6) group_sizes y group_names recalculados desde las columnas finales
    grp <- tibble::tibble(segment = seg_vec) %>% dplyr::count(segment, name = "n") %>% dplyr::arrange(segment)

    group_sizes <- as.integer(grp$n)
    group_names <- as.character(grp$segment)

    # validación final
    if (sum(group_sizes) != ncol(X_all)) return(NULL)
    if (any(group_sizes <= 0)) return(NULL)

    list(
      respondent_group = respondent_group,
      instrument = instrument,
      X = as.data.frame(X_all),
      meta = meta_all,
      group_sizes = group_sizes,
      group_names = group_names,
      seg_vec = seg_vec
    )
  })

  out <- out[!vapply(out, is.null, logical(1))]
  out
}


```

```{r}
run_mfa <- function(X, group_sizes, group_names,
                    ncp = 5, impute = TRUE, graph = FALSE,
                    min_rows = 3, min_cols = 2) {

  X <- as.data.frame(X)

  # ---- Validaciones duras ----
  if (nrow(X) < min_rows) stop("MFA: too few rows.")
  if (ncol(X) < min_cols) stop("MFA: too few columns.")
  if (length(group_sizes) == 0) stop("MFA: group_sizes is empty.")
  if (any(group_sizes <= 0)) stop("MFA: some group_sizes <= 0.")
  if (sum(group_sizes) != ncol(X)) stop("MFA: sum(group_sizes) != ncol(X).")

  has_na <- anyNA(as.matrix(X))

  # ---- Imputación SOLO si hay NA ----
  if (impute && has_na) {
    imp <- missMDA::imputeMFA(
      X,
      group = group_sizes,
      type  = rep("s", length(group_sizes)),
      ncp   = ncp
    )
    X_use <- imp$completeObs
  } else {
    X_use <- X
  }

  FactoMineR::MFA(
    X_use,
    group      = group_sizes,
    type       = rep("s", length(group_sizes)),
    name.group = group_names,
    ncp        = ncp,
    graph      = graph
  )
}

```

```{r}
combo_list <- mfa_inputs_by_combo(survey_db, keep_ds = TRUE, drop_constant = TRUE)

mfa_results <- purrr::map(combo_list, function(obj) {
  key <- paste(obj$respondent_group, obj$instrument, sep = " | ")

  res <- tryCatch(
    run_mfa(obj$X, obj$group_sizes, obj$group_names, ncp = 5, impute = TRUE, graph = FALSE),
    error = function(e) e
  )

  if (inherits(res, "error")) {
    message("MFA FAILED: ", key, " -> ", res$message)
    return(list(meta = obj[c("respondent_group","instrument")], ok = FALSE, error = res$message))
  }

  list(meta = obj[c("respondent_group","instrument")], ok = TRUE, res = res, obj = obj)
})

# qué combos quedaron OK
purrr::map_chr(mfa_results, \(z) paste0(z$meta$respondent_group, " | ", z$meta$instrument, " | ok=", z$ok))

```

```{r}
ok_ids <- which(vapply(mfa_results, \(z) isTRUE(z$ok), logical(1)))
one <- mfa_results[[ ok_ids[1] ]]$res

factoextra::fviz_screeplot(one)
factoextra::fviz_mfa_ind(one, repel = TRUE)
factoextra::fviz_mfa_var(one, repel = TRUE)
factoextra::fviz_mfa_group(one, repel = TRUE)

```

#####################################################################################
#####################################################################################
#####################################################################################
#####################################################################################
#####################################################################################
#####################################################################################










```{r}
run_mfa_global_safe <- function(inp, ncp = 5, impute = TRUE, graph = FALSE) {

  X <- as.data.frame(inp$X)

  # Imputa solo si hay NA
  if (impute && anyNA(as.matrix(X))) {
    imp <- missMDA::imputeMFA(
      X,
      group = inp$group_sizes,
      type  = rep("s", length(inp$group_sizes)),
      ncp   = ncp
    )
    X_use <- as.data.frame(imp$completeObs)
  } else {
    X_use <- X
  }

  mfa <- FactoMineR::MFA(
    X_use,
    group      = inp$group_sizes,
    type       = rep("s", length(inp$group_sizes)),
    name.group = inp$group_names,
    ncp        = ncp,
    graph      = graph
  )

  # guardo metadatos como atributo para usarlos luego
  attr(mfa, "meta") <- inp$meta
  mfa
}

```

```{r}
mfa_input_global <- function(survey_db,
                             keep_ds = TRUE,
                             drop_constant = TRUE,
                             min_rows = 3,
                             min_cols = 2) {

  # 1) Extraer X y meta de cada hoja
  inputs <- purrr::imap(survey_db, function(sheet_obj, sheet_name) {

    df      <- sheet_obj$data
    seg_map <- sheet_obj$seg_map

    item_cols <- get_item_columns(df, seg_map)
    if (length(item_cols) < 2) return(NULL)

    X <- df %>%
      dplyr::select(dplyr::all_of(item_cols)) %>%
      dplyr::mutate(dplyr::across(dplyr::everything(),
                                  ~ suppressWarnings(as.numeric(as.character(.x))))) %>%
      as.data.frame()

    meta <- df %>%
      dplyr::transmute(
        respondent_id = .data[[names(df)[1]]],
        ds = .data[[names(df)[2]]],
        respondent_group = .data[["respondent_group"]],
        instrument = .data[["instrument"]],
        sheet = sheet_name
      )

    if (!keep_ds) meta <- meta %>% dplyr::select(-ds)

    list(sheet = sheet_name, X = X, meta = meta)
  })

  inputs <- inputs[!vapply(inputs, is.null, logical(1))]
  if (!length(inputs)) stop("No hay hojas válidas para construir el MFA global.")

  # 2) Union de columnas (todos los items que existan en cualquier hoja)
  all_items <- Reduce(union, lapply(inputs, \(z) colnames(z$X)))

  # 3) REVISION: link the instruments of the SAME respondent horizontally (one row per participant).
  #    The submitted code stacked the six sheets by rows (purrr::map_dfr), which turned 77 respondents into 231
  #    instrument-level pseudo-participants and created 7,854 structural NAs that were later mode-imputed.
  # --- submitted code (disabled) ---
  # X_all <- purrr::map_dfr(inputs, \(z){ X <- z$X; miss <- setdiff(all_items, colnames(X)); if (length(miss)) X[miss] <- NA_real_; X[, all_items, drop = FALSE] })
  # meta_all <- purrr::map_dfr(inputs, \(z) z$meta)
  by_group <- split(inputs, vapply(inputs, \(z) unique(as.character(z$meta$respondent_group)), character(1)))
  joined <- purrr::map(by_group, function(lst) {
    parts <- purrr::map(lst, \(z) dplyr::bind_cols(dplyr::select(z$meta, respondent_id, ds, respondent_group), z$X))
    purrr::reduce(parts, dplyr::inner_join, by = c("respondent_id", "ds", "respondent_group"), relationship = "one-to-one")
  })
  n_in  <- vapply(by_group, \(lst) nrow(lst[[1]]$X), integer(1)); n_out <- vapply(joined, nrow, integer(1))
  n_all <- vapply(by_group, \(lst) length(unique(unlist(lapply(lst, \(z) nrow(z$X))))), integer(1))
  if (any(n_all != 1) || !identical(unname(n_in), unname(n_out))) stop("REVISION check: participants differ across instruments or were lost/duplicated in the join.")
  wide <- dplyr::bind_rows(joined)      # students and educators are different people: binding the groups by rows is correct
  if (anyDuplicated(wide[, c("respondent_group", "respondent_id")])) stop("REVISION check: duplicated participant after join.")
  X_all <- as.data.frame(wide[, all_items, drop = FALSE])
  meta_all <- dplyr::transmute(wide, respondent_id, ds, respondent_group, instrument = "CCI + HxC + TEI", sheet = "participant-level")
  if (anyNA(as.matrix(X_all))) stop("REVISION check: missing values in the participant-level matrix (no imputation is performed).")

  # 4) Limpieza columnas
  # (a) quitar all-NA
  keep_not_allna <- colSums(!is.na(X_all)) > 0
  X_all <- X_all[, keep_not_allna, drop = FALSE]

  if (nrow(X_all) < min_rows || ncol(X_all) < min_cols) stop("Muy pocos datos tras all-NA.")

  # (b) quitar var=0 si quieres
  if (drop_constant) {
    v <- vapply(X_all, function(z) var(z, na.rm = TRUE), numeric(1))
    keep_var <- is.finite(v) & v > 0
    X_all <- X_all[, keep_var, drop = FALSE]
  }

  if (ncol(X_all) < min_cols) stop("Muy pocas columnas tras var=0.")

  # 5) Construir segmentación “global” por columna
  # Buscamos el segmento del item en cualquier hoja (primera que lo tenga)
  get_seg_global <- function(items, survey_db) {
    seg <- rep(NA_character_, length(items))
    names(seg) <- items

    for (sh in names(survey_db)) {
      sm <- survey_db[[sh]]$seg_map
      hit <- sm[items]
      to_fill <- is.na(seg) & !is.na(hit)
      seg[to_fill] <- unname(hit[to_fill])
      if (!any(is.na(seg))) break
    }
    seg[is.na(seg) | stringr::str_trim(seg) == ""] <- "Unsegmented"
    seg
  }

  seg_vec <- get_seg_global(colnames(X_all), survey_db)

  # (c) quitar Unsegmented
  keep_seg <- stringr::str_trim(seg_vec) != "Unsegmented"
  X_all <- X_all[, keep_seg, drop = FALSE]
  seg_vec <- seg_vec[keep_seg]

  if (ncol(X_all) < min_cols) stop("Muy pocas columnas tras Unsegmented.")

  # 6) Ordenar por segmento
  ord <- order(seg_vec, colnames(X_all))
  X_all <- X_all[, ord, drop = FALSE]
  seg_vec <- seg_vec[ord]

  # 7) group_sizes / group_names
  grp <- tibble::tibble(segment = seg_vec) %>%
    dplyr::count(segment, name = "n") %>%
    dplyr::arrange(segment)

  group_sizes <- as.integer(grp$n)
  group_names <- as.character(grp$segment)

  list(
    X = as.data.frame(X_all),
    meta = meta_all,
    group_sizes = group_sizes,
    group_names = group_names,
    seg_vec = seg_vec
  )
}

inp_global <- mfa_input_global(survey_db, keep_ds = TRUE, drop_constant = TRUE)
mfa_global <- run_mfa_global_safe(inp_global, ncp = 5, impute = TRUE)

meta <- attr(mfa_global, "meta")

# Scree
factoextra::fviz_screeplot(mfa_global, addlabels = TRUE)

# Individuos coloreados por DS
factoextra::fviz_mfa_ind(
  mfa_global,
  habillage = as.factor(meta$ds),
  addEllipses = TRUE,
  repel = TRUE
)

# Individuos coloreados por grupo
factoextra::fviz_mfa_ind(
  mfa_global,
  habillage = as.factor(meta$respondent_group),
  addEllipses = TRUE,
  repel = TRUE
)

# Individuos coloreados por instrumento
# REVISION 2026: plot removed - with one row per participant, 'instrument' is no longer a row attribute
# factoextra::fviz_mfa_ind(mfa_global, habillage = as.factor(meta$instrument), addEllipses = TRUE, repel = TRUE)

# Variables (selecciona top contrib para que no sea ilegible)
factoextra::fviz_mfa_var(
  mfa_global,
  "quanti.var",
  repel = TRUE,
  select.var = list(contrib = 20)
)

# Grupos de variables
factoextra::fviz_mfa_group(mfa_global, repel = TRUE)

```

```{r}
coord <- as.data.frame(mfa_global$ind$coord)
meta  <- attr(mfa_global, "meta")

df_dim <- dplyr::bind_cols(meta, coord)

# ANOVA simple (ejemplo Dim1)
summary(aov(Dim.1 ~ respondent_group, data = df_dim))
# summary(aov(Dim.1 ~ instrument, data = df_dim))   # REVISION 2026: instrument is no longer a row attribute (one row per participant)
summary(aov(Dim.1 ~ ds, data = df_dim))

# Modelo conjunto (más serio)
summary(lm(Dim.1 ~ respondent_group + ds, data = df_dim))   # REVISION 2026: instrument term removed
summary(lm(Dim.2 ~ respondent_group + ds, data = df_dim))   # REVISION 2026: instrument term removed

```
















#####################################################################################
#####################################################################################
#####################################################################################
#####################################################################################
#####################################################################################
#####################################################################################






```{r}
pkgs <- c("dplyr","tibble","stringr","purrr","missMDA","FactoMineR","factoextra","psych")
to_install <- pkgs[!pkgs %in% rownames(installed.packages())]
if(length(to_install)) install.packages(to_install)
invisible(lapply(pkgs, library, character.only = TRUE))

```

```{r}
impute_mode_ordinal <- function(df) {
  out <- df
  for (j in seq_along(out)) {
    x <- out[[j]]
    if (all(is.na(x))) next
    tab <- table(x, useNA = "no")
    mode_val <- names(tab)[which.max(tab)]
    out[[j]][is.na(out[[j]])] <- mode_val
  }
  out
}

```

```{r}
poly_pca_scores <- function(df_block, ncp = 2) {
  # df_block: data.frame con ítems Likert (1-5) (idealmente como ordered)
  df_block <- as.data.frame(df_block)

  # Imputa NA (modo) para poder calcular policórica
  stopifnot(!anyNA(df_block))   # REVISION 2026: participant-level data contain no NA, so the mode imputation below is a no-op
  df_imp <- impute_mode_ordinal(df_block)

  # Asegura ordinal (ordered factor) si viene como num/char
  df_imp <- df_imp %>%
    mutate(across(everything(), ~ {
      if (is.factor(.x) && is.ordered(.x)) return(.x)
      # fuerza niveles 1..5 si aplica
      ordered(as.character(.x), levels = sort(unique(as.character(.x))))
    }))

  # Correlación policórica
  pc <- psych::polychoric(df_imp)
  R  <- pc$rho

  # PCA (eigen)
  eig <- eigen(R, symmetric = TRUE)
  loadings <- eig$vectors[, 1:ncp, drop = FALSE] %*% diag(sqrt(pmax(eig$values[1:ncp], 0)))
  colnames(loadings) <- paste0("PC", 1:ncp)
  rownames(loadings) <- colnames(df_imp)

  # Datos numéricos estandarizados (solo para construir scores)
  Z <- df_imp %>%
    mutate(across(everything(), ~ as.numeric(as.character(.x)))) %>%
    as.matrix()
  Z <- scale(Z)

  # Scores tipo regresión: W = R^{-1} L (L' R^{-1} L)^{-1}
  Rinv <- tryCatch(solve(R), error = function(e) MASS::ginv(R))
  mid  <- tryCatch(solve(t(loadings) %*% Rinv %*% loadings),
                   error = function(e) MASS::ginv(t(loadings) %*% Rinv %*% loadings))
  W <- Rinv %*% loadings %*% mid

  scores <- Z %*% W
  scores <- as.data.frame(scores)
  colnames(scores) <- paste0("PC", 1:ncp)

  # Varianza explicada (por eigenvalues / p)
  p <- ncol(R)
  var_exp <- eig$values[1:ncp] / sum(eig$values)
  names(var_exp) <- paste0("PC", 1:ncp)

  list(scores = scores, var_exp = var_exp, R = R, loadings = loadings)
}

```

```{r}
build_poly_scores_by_block <- function(inp, ncp_block = 2) {

  X <- inp$X
  seg_vec <- inp$seg_vec

  # split columnas por bloque
  blocks <- split(colnames(X), seg_vec)

  # para cada bloque => scores (n x ncp_block)
  score_list <- lapply(names(blocks), function(g) {
    cols <- blocks[[g]]
    res <- poly_pca_scores(X[, cols, drop = FALSE], ncp = ncp_block)
    out <- res$scores
    colnames(out) <- paste0(g, "__PC", seq_len(ncp_block))
    attr(out, "var_exp") <- res$var_exp
    out
  })
  names(score_list) <- names(blocks)

  scores_all <- dplyr::bind_cols(score_list)

  # grupos del nuevo MFA: cada bloque aporta ncp_block columnas
  group_sizes2 <- rep(ncp_block, length(blocks))
  group_names2 <- names(blocks)

  list(scores = as.data.frame(scores_all),
       group_sizes = group_sizes2,
       group_names = group_names2,
       meta = inp$meta)
}

```

```{r}
inp_poly <- build_poly_scores_by_block(inp_global, ncp_block = 2)

mfa_poly <- FactoMineR::MFA(
  inp_poly$scores,
  group      = inp_poly$group_sizes,
  type       = rep("s", length(inp_poly$group_sizes)),
  name.group = inp_poly$group_names,
  ncp        = 5,
  graph      = FALSE
)

factoextra::fviz_screeplot(mfa_poly, addlabels = TRUE)
factoextra::fviz_mfa_group(mfa_poly, repel = TRUE)

```

```{r}
meta <- inp_poly$meta

factoextra::fviz_mfa_ind(mfa_poly, habillage = as.factor(meta$ds),
                         addEllipses = TRUE, repel = TRUE)

factoextra::fviz_mfa_ind(mfa_poly, habillage = as.factor(meta$respondent_group),
                         addEllipses = TRUE, repel = TRUE)

# REVISION 2026: instrument-coloured plot removed (instrument is no longer a row attribute)
# factoextra::fviz_mfa_ind(mfa_poly, habillage = as.factor(meta$instrument), addEllipses = TRUE, repel = TRUE)

```

```{r}
# ============================================================
# 0) Paquetes extra (por si no los tienes)
# ============================================================
pkgs2 <- c("dplyr","tibble","purrr")
to_install2 <- pkgs2[!pkgs2 %in% rownames(installed.packages())]
if(length(to_install2)) install.packages(to_install2)
invisible(lapply(pkgs2, library, character.only = TRUE))

# ============================================================
# 1) Helper: % var explicada Dim1+Dim2 para un MFA
#    (FactoMineR guarda eigenvalues en $eig)
# ============================================================
mfa_dim12_explained <- function(mfa_obj) {
  stopifnot(!is.null(mfa_obj$eig))
  # en FactoMineR: columna 2 = porcentaje por dimensión
  d1 <- mfa_obj$eig[1, 2]
  d2 <- mfa_obj$eig[2, 2]
  tibble::tibble(
    Dim1 = d1,
    Dim2 = d2,
    Dim1_Dim2 = d1 + d2
  )
}

# ============================================================
# 2) Helper: tabla var_exp por bloque (PC1/PC2) + resumen
#    Usa la misma lógica de bloques (seg_vec) y vuelve a correr
#    poly_pca_scores() por bloque para recuperar var_exp.
# ============================================================
poly_var_table_by_block <- function(inp, ncp_block = 2) {
  X <- inp$X
  seg_vec <- inp$seg_vec
  blocks <- split(colnames(X), seg_vec)

  # recalcula policórico por bloque y guarda var_exp
  out <- purrr::map_dfr(names(blocks), function(g) {
    cols <- blocks[[g]]
    # OJO: poly_pca_scores debe estar definido en tu script
    res <- poly_pca_scores(X[, cols, drop = FALSE], ncp = ncp_block)

    ve <- res$var_exp
    # si algún bloque tiene menos de ncp_block posible, lo protegemos
    pc1 <- if(length(ve) >= 1) unname(ve[1]) else NA_real_
    pc2 <- if(length(ve) >= 2) unname(ve[2]) else NA_real_

    tibble::tibble(
      block = g,
      n_items = length(cols),
      PC1 = pc1,
      PC2 = pc2,
      PC1_PC2 = pc1 + pc2
    )
  })

  out %>%
    dplyr::arrange(dplyr::desc(PC1_PC2))
}

# ============================================================
# 3) Corre MFA policórico por bloques (si no lo tienes ya)
# ============================================================
# ncp_block = 2 => sacas PC1 y PC2 por bloque
inp_poly <- build_poly_scores_by_block(inp_global, ncp_block = 2)

mfa_poly <- FactoMineR::MFA(
  inp_poly$scores,
  group      = inp_poly$group_sizes,
  type       = rep("s", length(inp_poly$group_sizes)),
  name.group = inp_poly$group_names,
  ncp        = 5,
  graph      = FALSE
)

# ============================================================
# 4) Reporte comparativo "antes vs después"
# ============================================================
before <- mfa_dim12_explained(mfa_global) %>%
  dplyr::mutate(model = "MFA clásico (ítems originales)")

after  <- mfa_dim12_explained(mfa_poly) %>%
  dplyr::mutate(model = "MFA policórico por bloques (PCs)")

compare_dim12 <- dplyr::bind_rows(before, after) %>%
  dplyr::select(model, Dim1, Dim2, Dim1_Dim2)

print(compare_dim12)

# Mensaje amigable
cat("\n--- Resumen ---\n")
cat("Dim1+Dim2 (clásico):   ", round(before$Dim1_Dim2, 2), "%\n")
cat("Dim1+Dim2 (policórico):", round(after$Dim1_Dim2, 2), "%\n")
cat("Diferencia:            ", round(after$Dim1_Dim2 - before$Dim1_Dim2, 2), "p.p.\n")

# ============================================================
# 5) Tabla por bloque: var_exp PC1/PC2 (justificación)
# ============================================================
tab_blocks <- poly_var_table_by_block(inp_global, ncp_block = 2)

print(tab_blocks)

# Un resumen global (promedios simples)
cat("\n--- Bloques (promedio simple) ---\n")
cat("Promedio PC1:     ", round(mean(tab_blocks$PC1, na.rm = TRUE) * 100, 1), "%\n")
cat("Promedio PC2:     ", round(mean(tab_blocks$PC2, na.rm = TRUE) * 100, 1), "%\n")
cat("Promedio PC1+PC2: ", round(mean(tab_blocks$PC1_PC2, na.rm = TRUE) * 100, 1), "%\n")

```

```{r}
pkgs <- c("dplyr","tibble","stringr","purrr","missMDA","FactoMineR","factoextra","psych","MASS")
to_install <- pkgs[!pkgs %in% rownames(installed.packages())]
if(length(to_install)) install.packages(to_install)
invisible(lapply(pkgs, library, character.only = TRUE))


impute_mode_ordinal <- function(df) {
  out <- df
  for (j in seq_along(out)) {
    x <- out[[j]]
    if (all(is.na(x))) next
    tab <- table(x, useNA = "no")
    mode_val <- names(tab)[which.max(tab)]
    out[[j]][is.na(out[[j]])] <- mode_val
  }
  out
}

poly_pca_scores <- function(df_block, ncp = 2) {
  df_block <- as.data.frame(df_block)

  stopifnot(!anyNA(df_block))   # REVISION 2026: participant-level data contain no NA, so the mode imputation below is a no-op
  df_imp <- impute_mode_ordinal(df_block)

  df_imp <- df_imp %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), ~ {
      if (is.factor(.x) && is.ordered(.x)) return(.x)
      ordered(as.character(.x), levels = sort(unique(as.character(.x))))
    }))

  pc <- psych::polychoric(df_imp)
  R  <- pc$rho

  eig <- eigen(R, symmetric = TRUE)
  loadings <- eig$vectors[, 1:ncp, drop = FALSE] %*% diag(sqrt(pmax(eig$values[1:ncp], 0)))
  colnames(loadings) <- paste0("PC", 1:ncp)
  rownames(loadings) <- colnames(df_imp)

  Z <- df_imp %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), ~ as.numeric(as.character(.x)))) %>%
    as.matrix()
  Z <- scale(Z)

  Rinv <- tryCatch(solve(R), error = function(e) MASS::ginv(R))
  mid  <- tryCatch(solve(t(loadings) %*% Rinv %*% loadings),
                   error = function(e) MASS::ginv(t(loadings) %*% Rinv %*% loadings))
  W <- Rinv %*% loadings %*% mid

  scores <- Z %*% W
  scores <- as.data.frame(scores)
  colnames(scores) <- paste0("PC", 1:ncp)

  var_exp <- eig$values[1:ncp] / sum(eig$values)
  names(var_exp) <- paste0("PC", 1:ncp)

  list(scores = scores, var_exp = var_exp, R = R, loadings = loadings)
}

# ------------------------------------------------------------
# AUTOMÁTICO: 1 vs 2 PCs por bloque
# ------------------------------------------------------------
build_poly_scores_by_block_auto <- function(inp,
                                            max_pcs = 2,
                                            thr_pc1  = 0.85,
                                            thr_cum2 = 0.75,
                                            min_items_for_pca = 2) {

  X <- inp$X
  seg_vec <- inp$seg_vec
  blocks <- split(colnames(X), seg_vec)

  block_summ <- list()
  score_list  <- list()

  for (g in names(blocks)) {
    cols <- blocks[[g]]

    # Si el bloque tiene <2 ítems, no hay PCA real. Caso extremo:
    if (length(cols) < min_items_for_pca) {
      # alternativa simple: usar el ítem tal cual como "PC1"
      sc <- X[, cols, drop = FALSE]
      colnames(sc) <- paste0(g, "__PC1")
      score_list[[g]] <- as.data.frame(sc)
      block_summ[[g]] <- tibble::tibble(
        block = g, n_items = length(cols),
        pc_used = 1,
        PC1 = NA_real_, PC2 = NA_real_, PC1_PC2 = NA_real_,
        rule = "n_items<2 -> uso ítem como PC1"
      )
      next
    }

    # Calcula hasta 2 PCs (policórica + PCA)
    res2 <- poly_pca_scores(X[, cols, drop = FALSE], ncp = max_pcs)
    ve <- res2$var_exp
    pc1 <- unname(ve[1])
    pc2 <- if (length(ve) >= 2) unname(ve[2]) else NA_real_
    cum2 <- if (!is.na(pc2)) pc1 + pc2 else pc1

    # Decide pcs
    pcs_use <- if (!is.na(pc1) && pc1 >= thr_pc1) 1 else 2
    rule_txt <- if (pcs_use == 1) {
      paste0("PC1>= ", thr_pc1)
    } else if (!is.na(cum2) && cum2 >= thr_cum2) {
      paste0("PC1<", thr_pc1, " y PC1+PC2>= ", thr_cum2)
    } else {
      paste0("PC1+PC2<", thr_cum2, " -> uso 2 PCs por defecto")
    }

    sc <- res2$scores[, 1:pcs_use, drop = FALSE]
    colnames(sc) <- paste0(g, "__PC", seq_len(pcs_use))
    score_list[[g]] <- as.data.frame(sc)

    block_summ[[g]] <- tibble::tibble(
      block = g, n_items = length(cols),
      pc_used = pcs_use,
      PC1 = pc1,
      PC2 = pc2,
      PC1_PC2 = cum2,
      rule = rule_txt
    )
  }

  scores_all <- dplyr::bind_cols(score_list)

  # group sizes = pcs usados por bloque
  summ_df <- dplyr::bind_rows(block_summ) %>% dplyr::arrange(block)
  group_sizes2 <- summ_df$pc_used
  group_names2 <- summ_df$block

  list(
    scores = as.data.frame(scores_all),
    group_sizes = as.integer(group_sizes2),
    group_names = as.character(group_names2),
    block_summary = summ_df,
    meta = inp$meta
  )
}

# ------------------------------------------------------------
# RUN: MFA policórico auto (1/2 PCs por bloque) + reporte
# ------------------------------------------------------------
run_mfa_poly_auto <- function(inp_global,
                              thr_pc1 = 0.85,
                              thr_cum2 = 0.75,
                              ncp_mfa = 5,
                              graph = FALSE) {

  inp_auto <- build_poly_scores_by_block_auto(
    inp_global,
    max_pcs = 2,
    thr_pc1 = thr_pc1,
    thr_cum2 = thr_cum2
  )

  mfa_auto <- FactoMineR::MFA(
    inp_auto$scores,
    group      = inp_auto$group_sizes,
    type       = rep("s", length(inp_auto$group_sizes)),
    name.group = inp_auto$group_names,
    ncp        = ncp_mfa,
    graph      = graph
  )

  attr(mfa_auto, "meta") <- inp_auto$meta
  list(inp_auto = inp_auto, mfa_auto = mfa_auto)
}

# --------------------------
# Ejemplo de uso:
# res_auto <- run_mfa_poly_auto(inp_global, thr_pc1 = 0.85, thr_cum2 = 0.75)
# mfa_auto <- res_auto$mfa_auto
# res_auto$inp_auto$block_summary
# factoextra::fviz_screeplot(mfa_auto, addlabels = TRUE)
# meta <- attr(mfa_auto, "meta")
# factoextra::fviz_mfa_ind(mfa_auto, habillage = as.factor(meta$respondent_group), addEllipses = TRUE, repel=TRUE)

```

```{r message=FALSE, warning=FALSE}
pkgs <- c("dplyr","tibble","stringr","purrr","missMDA","FactoMineR","factoextra","psych","MASS","ggplot2","knitr")
to_install <- pkgs[!pkgs %in% rownames(installed.packages())]
if(length(to_install)) install.packages(to_install)
invisible(lapply(pkgs, library, character.only = TRUE))

dir.create(file.path(proj_dir, "archive", "previous_analysis", "outputs"), showWarnings = FALSE, recursive = TRUE)

# -----------------------------
# Helpers: tablas + var Dim1/2
# -----------------------------
mfa_dim12 <- function(mfa){
  # FactoMineR guarda %var en columna 2 de mfa$eig
  d1 <- as.numeric(mfa$eig[1,2])
  d2 <- as.numeric(mfa$eig[2,2])
  tibble::tibble(Dim1 = d1, Dim2 = d2, Dim1_Dim2 = d1 + d2)
}

save_table_csv <- function(df, filename){
  archive_outputs <- file.path(proj_dir, "archive", "previous_analysis", "outputs")
  dir.create(archive_outputs, showWarnings = FALSE, recursive = TRUE)
  path <- file.path(archive_outputs, filename)
  write.csv(df, path, row.names = FALSE)
  path
}

save_plot_pdf_png <- function(p, stem, w=10, h=7){
  archive_outputs <- file.path(proj_dir, "archive", "previous_analysis", "outputs")
  dir.create(archive_outputs, showWarnings = FALSE, recursive = TRUE)
  pdf_path <- file.path(archive_outputs, paste0(stem, ".pdf"))
  png_path <- file.path(archive_outputs, paste0(stem, ".png"))
  ggplot2::ggsave(pdf_path, plot = p, width = w, height = h, device = "pdf")
  ggplot2::ggsave(png_path, plot = p, width = w, height = h, dpi = 200)
  list(pdf = pdf_path, png = png_path)
}

# -----------------------------
# MFA clásico (global)
# -----------------------------
run_mfa_classic_global <- function(inp, ncp = 5, impute = TRUE, graph = FALSE){
  X <- as.data.frame(inp$X)

  if (impute && anyNA(as.matrix(X))) {
    imp <- missMDA::imputeMFA(
      X,
      group = inp$group_sizes,
      type  = rep("s", length(inp$group_sizes)),
      ncp   = ncp
    )
    X_use <- imp$completeObs
  } else {
    X_use <- X
  }

  mfa <- FactoMineR::MFA(
    X_use,
    group      = inp$group_sizes,
    type       = rep("s", length(inp$group_sizes)),
    name.group = inp$group_names,
    ncp        = ncp,
    graph      = graph
  )
  attr(mfa, "meta") <- inp$meta
  mfa
}

# -----------------------------
# Policórico por bloque: utilidades
# -----------------------------
impute_mode_ordinal <- function(df) {
  out <- df
  for (j in seq_along(out)) {
    x <- out[[j]]
    if (all(is.na(x))) next
    tab <- table(x, useNA = "no")
    mode_val <- names(tab)[which.max(tab)]
    out[[j]][is.na(out[[j]])] <- mode_val
  }
  out
}

poly_pca_scores <- function(df_block, ncp = 2) {
  df_block <- as.data.frame(df_block)
  stopifnot(!anyNA(df_block))   # REVISION 2026: participant-level data contain no NA, so the mode imputation below is a no-op
  df_imp <- impute_mode_ordinal(df_block)

  df_imp <- df_imp %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), ~ {
      if (is.factor(.x) && is.ordered(.x)) return(.x)
      ordered(as.character(.x), levels = sort(unique(as.character(.x))))
    }))

  pc <- psych::polychoric(df_imp)
  R  <- pc$rho

  eig <- eigen(R, symmetric = TRUE)
  loadings <- eig$vectors[, 1:ncp, drop = FALSE] %*% diag(sqrt(pmax(eig$values[1:ncp], 0)))
  colnames(loadings) <- paste0("PC", 1:ncp)

  Z <- df_imp %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), ~ as.numeric(as.character(.x)))) %>%
    as.matrix()
  Z <- scale(Z)

  Rinv <- tryCatch(solve(R), error = function(e) MASS::ginv(R))
  mid  <- tryCatch(solve(t(loadings) %*% Rinv %*% loadings),
                   error = function(e) MASS::ginv(t(loadings) %*% Rinv %*% loadings))
  W <- Rinv %*% loadings %*% mid

  scores <- as.data.frame(Z %*% W)
  colnames(scores) <- paste0("PC", 1:ncp)

  var_exp <- eig$values[1:ncp] / sum(eig$values)
  names(var_exp) <- paste0("PC", 1:ncp)

  list(scores = scores, var_exp = var_exp)
}

build_poly_scores_by_block_fixed <- function(inp, ncp_block = 2) {
  X <- inp$X
  seg_vec <- inp$seg_vec
  blocks <- split(colnames(X), seg_vec)

  score_list <- list()
  block_summary <- list()

  for (g in names(blocks)) {
    cols <- blocks[[g]]
    res <- poly_pca_scores(X[, cols, drop = FALSE], ncp = ncp_block)
    out <- res$scores
    colnames(out) <- paste0(g, "__PC", seq_len(ncp_block))
    score_list[[g]] <- out

    ve <- res$var_exp
    pc1 <- unname(ve[1]); pc2 <- unname(ve[2]); cum2 <- pc1 + pc2
    block_summary[[g]] <- tibble::tibble(
      block = g, n_items = length(cols),
      pc_used = ncp_block,
      PC1 = pc1, PC2 = pc2, PC1_PC2 = cum2,
      rule = paste0("fijo ", ncp_block, " PCs")
    )
  }

  scores_all <- dplyr::bind_cols(score_list)
  group_sizes2 <- rep(ncp_block, length(score_list))
  group_names2 <- names(score_list)

  list(
    scores = as.data.frame(scores_all),
    group_sizes = as.integer(group_sizes2),
    group_names = as.character(group_names2),
    block_summary = dplyr::bind_rows(block_summary) %>% dplyr::arrange(block),
    meta = inp$meta
  )
}

build_poly_scores_by_block_auto <- function(inp,
                                            thr_pc1  = 0.85,
                                            thr_cum2 = 0.75,
                                            min_items_for_pca = 2) {
  X <- inp$X
  seg_vec <- inp$seg_vec
  blocks <- split(colnames(X), seg_vec)

  score_list <- list()
  block_summary <- list()

  for (g in names(blocks)) {
    cols <- blocks[[g]]

    if (length(cols) < min_items_for_pca) {
      sc <- X[, cols, drop = FALSE]
      colnames(sc) <- paste0(g, "__PC1")
      score_list[[g]] <- as.data.frame(sc)

      block_summary[[g]] <- tibble::tibble(
        block = g, n_items = length(cols),
        pc_used = 1,
        PC1 = NA_real_, PC2 = NA_real_, PC1_PC2 = NA_real_,
        rule = "n_items<2 -> uso ítem como PC1"
      )
      next
    }

    res2 <- poly_pca_scores(X[, cols, drop = FALSE], ncp = 2)
    ve <- res2$var_exp
    pc1 <- unname(ve[1]); pc2 <- unname(ve[2]); cum2 <- pc1 + pc2

    pcs_use <- if (!is.na(pc1) && pc1 >= thr_pc1) 1 else 2
    rule_txt <- if (pcs_use == 1) {
      paste0("PC1>= ", thr_pc1)
    } else if (!is.na(cum2) && cum2 >= thr_cum2) {
      paste0("PC1<", thr_pc1, " y PC1+PC2>= ", thr_cum2)
    } else {
      paste0("PC1+PC2<", thr_cum2, " -> uso 2 PCs por defecto")
    }

    sc <- res2$scores[, 1:pcs_use, drop = FALSE]
    colnames(sc) <- paste0(g, "__PC", seq_len(pcs_use))
    score_list[[g]] <- as.data.frame(sc)

    block_summary[[g]] <- tibble::tibble(
      block = g, n_items = length(cols),
      pc_used = pcs_use,
      PC1 = pc1, PC2 = pc2, PC1_PC2 = cum2,
      rule = rule_txt
    )
  }

  scores_all <- dplyr::bind_cols(score_list)
  summ_df <- dplyr::bind_rows(block_summary) %>% dplyr::arrange(block)

  list(
    scores = as.data.frame(scores_all),
    group_sizes = as.integer(summ_df$pc_used),
    group_names = as.character(summ_df$block),
    block_summary = summ_df,
    meta = inp$meta
  )
}

run_mfa_from_scores <- function(score_obj, ncp = 5, graph = FALSE){
  mfa <- FactoMineR::MFA(
    score_obj$scores,
    group      = score_obj$group_sizes,
    type       = rep("s", length(score_obj$group_sizes)),
    name.group = score_obj$group_names,
    ncp        = ncp,
    graph      = graph
  )
  attr(mfa, "meta") <- score_obj$meta
  mfa
}

# -----------------------------
# Resumen narrativo por bloque
# -----------------------------
narrative_by_block <- function(block_df){
  block_df %>%
    dplyr::mutate(
      style = dplyr::case_when(
        pc_used == 1 ~ "unidimensional",
        pc_used == 2 & !is.na(PC2) & (PC2 / pmax(PC1, 1e-9)) < 0.25 ~ "bidimensional suave",
        pc_used == 2 ~ "bidimensional marcada",
        TRUE ~ "no definido"
      ),
      narrative = dplyr::case_when(
        pc_used == 1 ~ paste0(block, " unidimensional (PC1 captura casi toda la varianza intrabloque)."),
        style == "bidimensional suave" ~ paste0(block, " bidimensional suave (PC2 aporta señal secundaria)."),
        style == "bidimensional marcada" ~ paste0(block, " bidimensional marcada (PC2 aporta estructura relevante)."),
        TRUE ~ paste0(block, ": revisar.")
      )
    )
}

# ============================================================
# 1) CORRE LOS 3 MODELOS
# ============================================================
mfa_classic <- run_mfa_classic_global(inp_global, ncp = 5, impute = TRUE)
poly_fixed  <- build_poly_scores_by_block_fixed(inp_global, ncp_block = 2)
mfa_fixed   <- run_mfa_from_scores(poly_fixed, ncp = 5)

poly_auto   <- build_poly_scores_by_block_auto(inp_global, thr_pc1 = 0.85, thr_cum2 = 0.75)
mfa_auto    <- run_mfa_from_scores(poly_auto, ncp = 5)

meta_classic <- attr(mfa_classic, "meta")
meta_fixed   <- attr(mfa_fixed, "meta")
meta_auto    <- attr(mfa_auto, "meta")

# ============================================================
# 2) TABLA COMPARATIVA Dim1+Dim2 (HTML + CSV)
# ============================================================
cmp <- dplyr::bind_rows(
  dplyr::bind_cols(tibble::tibble(model="MFA clásico (ítems originales)"), mfa_dim12(mfa_classic)),
  dplyr::bind_cols(tibble::tibble(model="MFA policórico 2PC fijo/bloque"), mfa_dim12(mfa_fixed)),
  dplyr::bind_cols(tibble::tibble(model="MFA policórico AUTO (1 vs 2 PCs/bloque)"), mfa_dim12(mfa_auto))
)

cmp_path <- save_table_csv(cmp, "cmp_dim12_models.csv")
knitr::kable(cmp, digits = 2, caption = "Comparación de varianza explicada por Dim1/Dim2 (tres modelos)")

cat("\nArchivo guardado:", cmp_path, "\n")

# ============================================================
# 3) TABLA POR BLOQUE (AUTO) + NARRATIVA (HTML + CSV)
# ============================================================
block_tbl <- poly_auto$block_summary %>%
  dplyr::mutate(
    PC1 = round(PC1, 3),
    PC2 = round(PC2, 3),
    PC1_PC2 = round(PC1_PC2, 3)
  )

block_path <- save_table_csv(block_tbl, "block_summary_auto.csv")
knitr::kable(block_tbl, caption = "AUTO: Varianza intrabloque (PC1/PC2), PCs usados y regla de decisión")

cat("\nArchivo guardado:", block_path, "\n")

narr <- narrative_by_block(poly_auto$block_summary) %>%
  dplyr::select(block, n_items, pc_used, style, narrative)

narr_path <- save_table_csv(narr, "block_narrative_auto.csv")
knitr::kable(narr, caption = "AUTO: Mini-resumen narrativo por bloque")

cat("\nArchivo guardado:", narr_path, "\n")

# ============================================================
# 4) GRÁFICOS (se guardan PDF+PNG y se muestran PNG en HTML)
#    - “que me muestre todo”: scree + group + ind por DS/grupo/instrument
# ============================================================
make_plots <- function(mfa, meta, prefix){
  plots <- list(
    scree = factoextra::fviz_screeplot(mfa, addlabels = TRUE),
    groups = factoextra::fviz_mfa_group(mfa, repel = TRUE),
    ind_ds = factoextra::fviz_mfa_ind(mfa, habillage = as.factor(meta$ds), addEllipses = TRUE, repel = TRUE),
    ind_resp = factoextra::fviz_mfa_ind(mfa, habillage = as.factor(meta$respondent_group), addEllipses = TRUE, repel = TRUE)
    # REVISION 2026: ind_instr removed (instrument is no longer a row attribute)
  )

  out_files <- list()
  for (nm in names(plots)) {
    stem <- paste0(prefix, "_", nm)
    out_files[[nm]] <- save_plot_pdf_png(plots[[nm]], stem = stem, w = 11, h = 7)
  }
  out_files
}

files_classic <- make_plots(mfa_classic, meta_classic, "classic")
files_fixed   <- make_plots(mfa_fixed,   meta_fixed,   "poly_fixed2pc")
files_auto    <- make_plots(mfa_auto,    meta_auto,    "poly_auto")

# Mostrar en HTML (PNG)
show_pngs <- function(files_list, title){
  cat("\n\n### ", title, "\n\n")
  pngs <- unlist(lapply(files_list, function(x) x$png))
  knitr::include_graphics(pngs)
}

show_pngs(files_classic, "Gráficos — MFA clásico")
show_pngs(files_fixed,   "Gráficos — MFA policórico (2PC fijo por bloque)")
show_pngs(files_auto,    "Gráficos — MFA policórico AUTO (1 vs 2 PCs por bloque)")

cat("\n\n**Nota:** Los gráficos de este análisis anterior se conservan para trazabilidad en `archive/previous_analysis/outputs/`.\n")

```

```{r}
pkgs2 <- c("ggpubr")
to_install2 <- pkgs2[!pkgs2 %in% rownames(installed.packages())]
if(length(to_install2)) install.packages(to_install2)
invisible(lapply(pkgs2, library, character.only = TRUE))

```

```{r}
# ============================
# PANEL 2x2: MFA FINAL (elige cuál es tu "final": mfa_auto o mfa_fixed o mfa_classic)
# ============================

mfa_final  <- mfa_auto            # <- cambia a mfa_fixed o mfa_classic si ese es tu final
meta_final <- meta_auto           # <- idem: meta_fixed / meta_classic
prefix     <- "MFA_FINAL_panel_2x2"

# 4 plots (asegura mismos parámetros que ya venías usando)
p_scree  <- factoextra::fviz_screeplot(mfa_final, addlabels = TRUE)
p_group  <- factoextra::fviz_mfa_group(mfa_final, repel = TRUE)
p_var    <- fviz_mfa_ind(mfa_final, label = "none", habillage = as.factor(meta_final$respondent_group), addEllipses = TRUE, repel = TRUE)   # REVISION 2026: meta_final (was meta from an earlier chunk)

# individuos coloreados por DS (puedes cambiar a respondent_group/instrument si quieres)
p_ind    <- factoextra::fviz_mfa_ind(
  mfa_final,
  label = "none",
  habillage   = as.factor(meta_final$ds),
  addEllipses = TRUE,
  repel       = TRUE
)

# Panel 2x2 con etiquetas A–D
panel <- ggpubr::ggarrange(
  p_scree, p_group,
  p_var,   p_ind,
  ncol = 2, nrow = 2,
  labels = c("A", "B", "C", "D"),
  common.legend = FALSE
)

# Título opcional (se ve bien en paper)
#panel <- ggpubr::annotate_figure(
#  panel,
#  bottom = ggpubr::text_grob("Final MFA: Scree Plot, Groups, Variables, and #Individuals", face = "bold", size = 14)
#)

# Guardado paper-ready:
# - PDF vectorial (MEJOR para artículo)
# - Tamaño típico (A4 apaisado) o similar
archive_outputs <- file.path(proj_dir, "archive", "previous_analysis", "outputs")
dir.create(archive_outputs, showWarnings = FALSE, recursive = TRUE)

pdf_out <- file.path(archive_outputs, paste0(prefix, ".pdf"))
ggplot2::ggsave(pdf_out, plot = panel, device = cairo_pdf, width = 11.69, height = 8.27)

# Además (opcional) PNG 600 dpi por si la revista lo pide:
png_out <- file.path(archive_outputs, paste0(prefix, ".png"))
ggplot2::ggsave(png_out, plot = panel, width = 11.69, height = 8.27, dpi = 600)

cat("Guardado en:\n", pdf_out, "\n", png_out, "\n")

```




#####################################################################################
#####################################################################################
#####################################################################################

```{r}
build_famd_input_by_instrument <- function(survey_db, instrument_name,
                                           drop_constant = TRUE,
                                           min_rows = 5) {

  # seleccionar hojas del instrumento
  sel <- purrr::keep(survey_db, ~ .x$meta$instrument == instrument_name)
  if (length(sel) == 0) return(NULL)

  # unir datos
  df_all <- purrr::map_dfr(sel, ~ .x$data)

  # columnas base
  id_col <- names(df_all)[1]
  ds_col <- names(df_all)[2]

  # ítems segmentados
  seg_map <- sel[[1]]$seg_map
  item_cols <- get_item_columns(df_all, seg_map)

  if (length(item_cols) < 2) return(NULL)

  # dataset mixto
  df_mix <- df_all %>%
    dplyr::select(
      all_of(item_cols),
      ds,
      respondent_group
    ) %>%
    dplyr::mutate(
      dplyr::across(all_of(item_cols),
                    ~ suppressWarnings(as.numeric(as.character(.x)))),
      ds = as.factor(ds),
      respondent_group = as.factor(respondent_group)
    )

  # eliminar ítems constantes
  if (drop_constant) {
    v <- vapply(df_mix[item_cols], function(z) var(z, na.rm = TRUE), numeric(1))
    keep_items <- names(v)[is.finite(v) & v > 0]
    df_mix <- df_mix %>%
      dplyr::select(all_of(keep_items), ds, respondent_group)
  }

  if (nrow(df_mix) < min_rows || ncol(df_mix) < 3) return(NULL)

  df_mix
}

instruments <- c("CCI", "HxC", "TEI")

run_famd <- function(df_mix, ncp = 5, graph = FALSE) {

  FactoMineR::FAMD(
    df_mix,
    ncp = ncp,
    graph = graph
  )
}

famd_results <- purrr::map(instruments, function(inst) {

  df_mix <- build_famd_input_by_instrument(survey_db, inst)

  if (is.null(df_mix)) {
    message("FAMD skipped for ", inst)
    return(NULL)
  }

  res <- run_famd(df_mix, ncp = 5)

  list(
    instrument = inst,
    data = df_mix,
    res = res
  )
})

names(famd_results) <- instruments

```

```{r}
factoextra::fviz_famd_ind(
  famd_results$CCI$res,
  habillage = "respondent_group",
  addEllipses = TRUE,
  repel = TRUE
)

```

```{r}
factoextra::fviz_famd_ind(
  famd_results$CCI$res,
  habillage = "ds",
  addEllipses = TRUE,
  repel = TRUE
)

```

```{r}
factoextra::fviz_famd_var(
  famd_results$CCI$res,
  repel = TRUE
)

```

```{r}
factoextra::fviz_famd_ind(
  famd_results$HxC$res,
  habillage = "respondent_group",
  addEllipses = TRUE,
  repel = TRUE
)

```

```{r}
factoextra::fviz_famd_ind(
  famd_results$HxC$res,
  habillage = "ds",
  addEllipses = TRUE,
  repel = TRUE
)

```

```{r}
factoextra::fviz_famd_var(
  famd_results$HxC$res,
  repel = TRUE
)

```

```{r}
factoextra::fviz_famd_ind(
  famd_results$TEI$res,
  habillage = "respondent_group",
  addEllipses = TRUE,
  repel = TRUE
)

```

```{r}
factoextra::fviz_famd_ind(
  famd_results$TEI$res,
  habillage = "ds",
  addEllipses = TRUE,
  repel = TRUE
)

```

```{r}
factoextra::fviz_famd_var(
  famd_results$TEI$res,
  repel = TRUE
)

```

```{r}
# ============================================================
# Journal-ready FAMD plots (cos2 + contributions)
# ============================================================
pkgs <- c("factoextra","FactoMineR","ggplot2","dplyr","stringr","tibble")
to_install <- pkgs[!pkgs %in% rownames(installed.packages())]
if(length(to_install)) install.packages(to_install)
invisible(lapply(pkgs, library, character.only = TRUE))

# ---------- A) Style helpers ----------
theme_journal <- function(base_size = 12, base_family = "sans") {
  ggplot2::theme_minimal(base_size = base_size, base_family = base_family) +
    ggplot2::theme(
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(linewidth = 0.25),
      axis.title = element_text(face = "bold"),
      plot.title = element_text(face = "bold", size = base_size + 2),
      plot.subtitle = element_text(size = base_size),
      legend.title = element_text(face = "bold"),
      legend.position = "right"
    )
}

# Colorblind-friendly palette (Okabe-Ito)
pal_okabe_ito <- c(
  "#0072B2", "#D55E00", "#009E73", "#CC79A7",
  "#E69F00", "#56B4E9", "#F0E442", "#000000"
)

# ---------- B) Main function: plots for one instrument ----------
plot_famd_journal <- function(famd_res,
                              instrument = "CCI",
                              out_dir = file.path(proj_dir, "archive", "previous_analysis", "figures"),
                              dims = c(1,2),
                              ellipse = TRUE,
                              repel = TRUE,
                              label_points = "none",  # "none" recommended for journal; or "all"
                              dpi = 450) {

  dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

  # 1) Individuals: colored by respondent_group + cos2 as alpha/gradient
  p_ind_grp <- factoextra::fviz_famd_ind(
    famd_res,
    axes = dims,
    habillage = "respondent_group",
    addEllipses = ellipse,
    ellipse.level = 0.95,
    repel = repel,
    label = label_points
  ) +
    ggplot2::scale_color_manual(values = pal_okabe_ito) +
    ggplot2::labs(
      title = paste0("FAMD – Individuals (", instrument, ")"),
      subtitle = "Color = respondent_group (Students vs Educators), ellipses = 95% CI"
    ) +
    theme_journal()

  # 2) Individuals: quality (cos2) – shows which points are well represented on Dim1–Dim2
  #    This is useful to justify interpretation in paper.
  p_ind_cos2 <- factoextra::fviz_famd_ind(
    famd_res,
    axes = dims,
    col.ind = "cos2",      # gradient
    gradient.cols = c("#E6E6E6","#6BAED6","#2171B5","#08306B"),
    repel = repel,
    label = label_points
  ) +
    ggplot2::labs(
      title = paste0("FAMD – Individuals quality (cos²) (", instrument, ")"),
      subtitle = "Darker points = better represented on Dim1–Dim2"
    ) +
    theme_journal()

  # 3) Variables: contributions (quant + qual displayed together in FAMD)
  p_var_contrib <- factoextra::fviz_famd_var(
    famd_res,
    axes = dims,
    col.var = "contrib",
    gradient.cols = c("#F2F2F2","#FDBE85","#E6550D","#A63603"),
    repel = TRUE
  ) +
    ggplot2::labs(
      title = paste0("FAMD – Variables contribution (", instrument, ")"),
      subtitle = "Darker variables contribute more to Dim1–Dim2"
    ) +
    theme_journal()

  # 4) Save
  ggsave(file.path(out_dir, paste0("FAMD_", instrument, "_ind_group.png")),
         p_ind_grp, width = 7.2, height = 5.0, dpi = dpi, bg = "white")
  ggsave(file.path(out_dir, paste0("FAMD_", instrument, "_ind_cos2.png")),
         p_ind_cos2, width = 7.2, height = 5.0, dpi = dpi, bg = "white")
  ggsave(file.path(out_dir, paste0("FAMD_", instrument, "_var_contrib.png")),
         p_var_contrib, width = 7.2, height = 5.2, dpi = dpi, bg = "white")

  list(ind_group = p_ind_grp, ind_cos2 = p_ind_cos2, var_contrib = p_var_contrib)
}

# ---------- C) Tables: Top contributors (Dim1/Dim2) ----------
top_contrib_tables <- function(famd_res, top_n = 12) {

  # Quantitative vars contributions (items)
  q1 <- FactoMineR::dimdesc(famd_res, axes = 1:2, proba = 1)$Dim.1$quanti
  q2 <- FactoMineR::dimdesc(famd_res, axes = 1:2, proba = 1)$Dim.2$quanti

  # Robust fallback if dimdesc returns NULL (depending on data composition)
  # Use internal contrib matrices
  contrib_quanti <- tryCatch(famd_res$quanti.var$contrib, error = function(e) NULL)
  contrib_quali  <- tryCatch(famd_res$quali.var$contrib,  error = function(e) NULL)

  get_top <- function(mat, dim, top_n) {
    if (is.null(mat)) return(tibble::tibble(variable = character(), contrib = numeric()))
    tibble::tibble(variable = rownames(mat), contrib = mat[, dim]) %>%
      dplyr::arrange(dplyr::desc(contrib)) %>%
      dplyr::slice_head(n = top_n)
  }

  top_q_dim1 <- get_top(contrib_quanti, 1, top_n)
  top_q_dim2 <- get_top(contrib_quanti, 2, top_n)
  top_c_dim1 <- get_top(contrib_quali,  1, top_n)
  top_c_dim2 <- get_top(contrib_quali,  2, top_n)

  list(
    quanti_dim1 = top_q_dim1,
    quanti_dim2 = top_q_dim2,
    quali_dim1  = top_c_dim1,
    quali_dim2  = top_c_dim2
  )
}

# ============================================================
# RUN (example for CCI)
# ============================================================
# Pick your instrument result:
# famd_results$CCI$res  OR famd_results$HxC$res  OR famd_results$TEI$res

plots_CCI <- plot_famd_journal(
  famd_res = famd_results$CCI$res,
  instrument = "CCI",
  out_dir = file.path(proj_dir, "archive", "previous_analysis", "figures"),
  dims = c(1,2),
  ellipse = TRUE,
  repel = TRUE,
  label_points = "none",
  dpi = 450
)

top_CCI <- top_contrib_tables(famd_results$CCI$res, top_n = 12)

# Print tables in console (or knitr::kable in Rmd)
top_CCI$quanti_dim1
top_CCI$quanti_dim2
top_CCI$quali_dim1
top_CCI$quali_dim2

```

```{r}
plots_HxC <- plot_famd_journal(famd_results$HxC$res, instrument="HxC")
top_HxC   <- top_contrib_tables(famd_results$HxC$res)

plots_TEI <- plot_famd_journal(famd_results$TEI$res, instrument="TEI")
top_TEI   <- top_contrib_tables(famd_results$TEI$res)

```

```{r}
plots_CCI

```

```{r}
plots_HxC

```

```{r}
plots_TEI

```

#####################################################################################
# REVISION - ADDED SECTIONS (major revision, 2026-09)
#####################################################################################

The sections below were added during the revision. They (i) regenerate the manuscript tables from the objects computed above, (ii) reproduce the adherence figures from the source audit sheet, (iii) run a limited set of sensitivity analyses and an independent cross-check of the MFA, (iv) collect the evidence used to interpret the MFA axes, and (v) export everything used by the manuscript and response Rmds.

```{r rev-helpers}
# REVISION 2026: helpers
all_modes <- function(x) { t <- table(x); paste(sort(as.numeric(names(t)[t == max(t)])), collapse = "/") }
code_of <- setNames(item_map$code, item_map$name)
ins_lab <- c(CCI = "CCI", HxC = "D-OHR", TEI = "TEI")
```

```{r rev-table1}
# REVISION 2026: Table 1 (demographics) from the verified participant file (previously computed in Excel from the 'Demográfico' sheet,
# which contained excluded student #45, omitted #59 and assigned #51 to another school)
dem <- PARTICIPANTS %>% dplyr::mutate(ds = factor(ds, levels = paste0("DS", 1:7)))
T1_ds <- dem %>% dplyr::group_by(respondent_group, ds) %>% dplyr::summarise(n = dplyr::n(), female = sum(sex == "Femenino"), male = sum(sex == "Masculino"),
            age_mean = mean(age), age_sd = sd(age), .groups = "drop") %>% dplyr::mutate(female_pct = 100 * female / n, male_pct = 100 * male / n)
T1_tot <- dem %>% dplyr::group_by(respondent_group) %>% dplyr::summarise(n = dplyr::n(), female = sum(sex == "Femenino"), male = sum(sex == "Masculino"), age_mean = mean(age), age_sd = sd(age),
            age_min = min(age), age_max = max(age), .groups = "drop") %>% dplyr::mutate(female_pct = 100 * female / n, male_pct = 100 * male / n)
T1_perds <- dem %>% dplyr::count(respondent_group, ds) %>% dplyr::group_by(respondent_group) %>% dplyr::summarise(mean_per_ds = mean(n), sd_per_ds = sd(n))
knitr::kable(T1_ds, digits = 1); knitr::kable(T1_tot, digits = 1); knitr::kable(T1_perds, digits = 2)
```

```{r rev-table2-3}
# REVISION 2026: Table 2 = Mode & IQR by respondent group x instrument x segment (object ex_resp_instr_segment of the original script)
#           + Cronbach alpha by segment (object alpha_segment of the original script)
T2 <- ex_resp_instr_segment %>% dplyr::filter(instrument %in% c("CCI", "HxC")) %>%
  dplyr::left_join(alpha_segment %>% dplyr::select(respondent_group, instrument, segment, alpha, k_items, dropped_zero_var), by = c("respondent_group", "instrument", "segment")) %>%
  dplyr::mutate(n_items = sapply(seq_len(dplyr::n()), function(i) sum(item_map$instrument == instrument[i] & item_map$segment == segment[i])),
                all_modes = purrr::pmap_chr(list(respondent_group, instrument, segment), function(g, ins, s) all_modes(all_long$value[all_long$respondent_group == g & all_long$instrument == ins & all_long$segment == s])),
                instrument = ins_lab[instrument], alpha = ifelse(n_items < 2, NA, alpha))
stopifnot(nrow(T2) == 24, all(T2$n_nonmiss == T2$n_items * ifelse(T2$respondent_group == "Students", sum(PARTICIPANTS$respondent_group == "Students"), sum(PARTICIPANTS$respondent_group == "Educators"))))
# REVISION 2026: item-level Mode/IQR (Supplementary tables 2-3 and Table 3)
ITEMS <- mode_iqr_any_grouping(all_long, c("respondent_group", "instrument", "segment"), include_item = TRUE) %>%
  dplyr::mutate(code = unname(code_of[item]), all_modes = purrr::map2_chr(respondent_group, item, function(g, it) all_modes(all_long$value[all_long$respondent_group == g & all_long$item == it])),
                tie = grepl("/", all_modes), mode_first_in_data_order = mode, freq = purrr::map2_chr(respondent_group, item, function(g, it) paste(tabulate(all_long$value[all_long$respondent_group == g & all_long$item == it], 5), collapse = "/")))
T3_pooled <- ex_resp_instr_segment %>% dplyr::filter(instrument == "TEI") %>% dplyr::left_join(alpha_segment %>% dplyr::filter(instrument == "TEI") %>% dplyr::select(respondent_group, alpha), by = "respondent_group")
knitr::kable(T2, digits = 3); knitr::kable(ITEMS %>% dplyr::filter(tie), caption = "Items with tied modes (the original mode_numeric() returns the first tied value in data order)")
knitr::kable(T3_pooled, digits = 3)
# mean item score per participant, CCI vs D-OHR (descriptive, used for the D-OHR statement)
INS_MEANS <- PARTICIPANTS %>% dplyr::transmute(respondent_group, CCI = rowMeans(dplyr::across(dplyr::all_of(item_map$code[item_map$instrument == "CCI"]))),
                                               DOHR = rowMeans(dplyr::across(dplyr::all_of(item_map$code[item_map$instrument == "HxC"])))) %>%
  dplyr::group_by(respondent_group) %>% dplyr::summarise(CCI_mean = mean(CCI), DOHR_mean = mean(DOHR), pct_CCI_higher = 100 * mean(CCI > DOHR))
knitr::kable(INS_MEANS, digits = 2)
```

```{r rev-adherence}
# REVISION 2026: adherence reproduced from the source audit sheet (previously computed in Excel)
ADH <- ADH_SRC   # REVISION 2026: read once in the data-source chunk (raw sheet or de-identified public file)
stopifnot(setequal(ADH$student, as.integer(PARTICIPANTS_RAW$n[PARTICIPANTS_RAW$respondent_group == "Students"])))   # REVISION 2026: adherence sheet lists the 61 RAW students
ADH_sel <- ADH %>% dplyr::filter(n_patients > 0); ADH_ok <- ADH_sel %>% dplyr::filter(!not_completed)
stopifnot(setequal(ADH_sel$student, as.integer(PARTICIPANTS$n[PARTICIPANTS$respondent_group == "Students"])))   # one D-OHR per completer
ADH_CRIT <- purrr::imap_dfr(c(recall = "Risk-appropriate recall interval", fluoride = "Risk-appropriate fluoride application", noc_active = "Non-operative care, initial active lesions",
                              monitor_inactive = "Active monitoring, initial inactive lesions"),
                            ~ tibble::tibble(criterion = .x, applicable = sum(!is.na(ADH_ok[[.y]])), adherent = sum(ADH_ok[[.y]], na.rm = TRUE))) %>% dplyr::mutate(pct = 100 * adherent / applicable)
ADH_FLOW <- tibble::tibble(stage = c("Students in RAW questionnaire sheet (incl. 1 without patients)", "Students with >=1 treated patient (one D-OHR selected)", "Selected D-OHRs completed/assessable",
                                     "Applicable criterion-level decisions", "Adherent decisions", "Patients treated (all students)"),
                           n = c(nrow(ADH), nrow(ADH_sel), nrow(ADH_ok), sum(ADH_CRIT$applicable), sum(ADH_CRIT$adherent), sum(ADH$n_patients)))
ADH_TOT <- list(pct_assessable = 100 * nrow(ADH_ok) / nrow(ADH_sel), pct_overall = 100 * sum(ADH_CRIT$adherent) / sum(ADH_CRIT$applicable),
                pct_risk_nonadh = 100 * (1 - sum(ADH_CRIT$adherent[1:2]) / sum(ADH_CRIT$applicable[1:2])), patients_by_ds = tapply(ADH$n_patients, ADH$ds, sum))
knitr::kable(ADH_FLOW); knitr::kable(ADH_CRIT, digits = 1)
```

```{r rev-blocks}
# REVISION 2026: block-retention table (auto rule of the original script) + positive-definiteness documentation
pd_info <- purrr::imap_dfr(split(colnames(inp_global$X), inp_global$seg_vec), function(cols, b) {
  x <- as.data.frame(lapply(inp_global$X[, cols, drop = FALSE], function(z) ordered(z)))
  R0 <- suppressWarnings(suppressMessages(psych::polychoric(x, smooth = FALSE)))$rho
  tibble::tibble(block = b, min_eigen_unsmoothed = min(eigen(R0, only.values = TRUE)$values), pd = min(eigen(R0, only.values = TRUE)$values) > 1e-8,
                 n_cci_items = sum(item_map$instrument[match(cols, item_map$name)] == "CCI"), sparse_cells_lt5 = sum(sapply(inp_global$X[, cols, drop = FALSE], function(z) { f <- tabulate(z, 5); sum(f > 0 & f < 5) })))
})
BLOCKS <- poly_auto$block_summary %>% dplyr::left_join(pd_info, by = "block") %>%
  dplyr::mutate(PC1_PC2 = PC1 + PC2, interpretation = dplyr::case_when(
    n_items == 2 & pc_used == 1 ~ sprintf("2-item block; PC1 >= 85%% means polychoric r = %.2f (strong association), not demonstrated unidimensionality", 2 * PC1 - 1),
    pc_used == 2 & PC1_PC2 < 0.75 ~ "PC1+PC2 < 75%: two components retained by the planned default",
    TRUE ~ "rule applied as planned"), smoothing = ifelse(pd, "none", "psych::polychoric smoothed the non-positive-definite matrix"))
knitr::kable(BLOCKS, digits = 3)
```

```{r rev-sensitivity}
# REVISION 2026: limited sensitivity analyses
# (a) students only - same pipeline (functions of the original script) on the student sheets
inp_stu <- mfa_input_global(survey_db[c("Students CCI", "Students HxC", "Students TEI")], keep_ds = TRUE, drop_constant = TRUE)
poly_stu <- build_poly_scores_by_block_auto(inp_stu, thr_pc1 = 0.85, thr_cum2 = 0.75); mfa_stu <- run_mfa_from_scores(poly_stu, ncp = 5)
# (b) nonparametric bootstrap of the inertia (participants resampled within respondent group)
B <- as.integer(params$boot_B)   # 200 for the reported analysis
boot <- purrr::map_dfr(seq_len(B), function(b) {
  idx <- c(sample(which(inp_global$meta$respondent_group == "Students"), replace = TRUE), sample(which(inp_global$meta$respondent_group == "Educators"), replace = TRUE))
  ib <- list(X = inp_global$X[idx, ], seg_vec = inp_global$seg_vec, meta = inp_global$meta[idx, ])
  if (any(vapply(ib$X, function(z) var(z) == 0, logical(1)))) return(NULL)
  r <- tryCatch(run_mfa_from_scores(build_poly_scores_by_block_auto(ib, thr_pc1 = 0.85, thr_cum2 = 0.75), ncp = 5), error = function(e) NULL)
  if (is.null(r)) return(NULL)
  tibble::tibble(b = b, Dim1 = r$eig[1, 2], Dim2 = r$eig[2, 2], Dim12 = r$eig[2, 3]) })
BOOT <- boot %>% dplyr::summarise(replicates = dplyr::n(), dplyr::across(c(Dim1, Dim2, Dim12), list(lo = ~ quantile(.x, 0.025), med = median, hi = ~ quantile(.x, 0.975))))
SENS <- tibble::tibble(model = c("Primary: participant-level, students + educators, auto 1-vs-2 PCs", "Students only, auto 1-vs-2 PCs", "Fixed 2 PCs per block",
                                 "Classical MFA on the original items (no polychoric step)"),
                       rows = c(nrow(inp_global$X), nrow(inp_stu$X), nrow(inp_global$X), nrow(inp_global$X)),
                       Dim1 = c(mfa_auto$eig[1, 2], mfa_stu$eig[1, 2], mfa_fixed$eig[1, 2], mfa_classic$eig[1, 2]),
                       Dim2 = c(mfa_auto$eig[2, 2], mfa_stu$eig[2, 2], mfa_fixed$eig[2, 2], mfa_classic$eig[2, 2]),
                       Dim1_Dim2 = c(mfa_auto$eig[2, 3], mfa_stu$eig[2, 3], mfa_fixed$eig[2, 3], mfa_classic$eig[2, 3]),
                       r_Dim1_with_primary = c(1, abs(cor(mfa_auto$ind$coord[inp_global$meta$respondent_group == "Students", 1], mfa_stu$ind$coord[, 1])),
                                               abs(cor(mfa_auto$ind$coord[, 1], mfa_fixed$ind$coord[, 1])), abs(cor(mfa_auto$ind$coord[, 1], mfa_classic$ind$coord[, 1]))),
                       r_Dim2_with_primary = c(1, abs(cor(mfa_auto$ind$coord[inp_global$meta$respondent_group == "Students", 2], mfa_stu$ind$coord[, 2])),
                                               abs(cor(mfa_auto$ind$coord[, 2], mfa_fixed$ind$coord[, 2])), abs(cor(mfa_auto$ind$coord[, 2], mfa_classic$ind$coord[, 2]))))
knitr::kable(SENS, digits = 3); knitr::kable(BOOT, digits = 1)
```

```{r rev-crosscheck}
# REVISION 2026: independent cross-check of the primary MFA with a second implementation
# polycor::polychor full maximum likelihood (pairwise) instead of the two-step psych::polychoric estimator, PCA scores by eigen-decomposition, and MFA computed
# directly by SVD (each block's standardized component scores weighted by 1/sqrt(first eigenvalue of the block)) instead of FactoMineR.
xc_blocks <- split(colnames(inp_global$X), inp_global$seg_vec); xc_scores <- list()
for (b in names(xc_blocks)) {
  x <- inp_global$X[, xc_blocks[[b]], drop = FALSE]; k <- ncol(x); R <- diag(k)
  for (i in 1:(k - 1)) for (j in (i + 1):k) R[i, j] <- R[j, i] <- polycor::polychor(x[[i]], x[[j]], ML = TRUE)   # full ML (thresholds + correlation jointly); psych uses the two-step estimator
  e <- eigen(R, symmetric = TRUE); if (min(e$values) < 1e-8) { R <- psych::cor.smooth(R); e <- eigen(R, symmetric = TRUE) }
  pcs <- if (e$values[1] / sum(e$values) >= 0.85) 1 else 2
  xc_scores[[b]] <- scale(scale(as.matrix(x)) %*% e$vectors[, 1:pcs, drop = FALSE])
}
wblk <- lapply(xc_scores, function(S) S / sqrt(max(eigen(cor(S), only.values = TRUE)$values)))
Zall <- do.call(cbind, wblk) / sqrt(nrow(inp_global$X))                                      # FactoMineR uses 1/n row weights
sv <- svd(scale(Zall, scale = FALSE)); xc_pct <- 100 * sv$d^2 / sum(sv$d^2)
XCHECK <- tibble::tibble(implementation = c("FactoMineR::MFA on psych::polychoric scores (primary)", "Manual SVD on polycor::polychor (full ML) scores (cross-check)"),
                         Dim1 = c(mfa_auto$eig[1, 2], xc_pct[1]), Dim2 = c(mfa_auto$eig[2, 2], xc_pct[2]), Dim1_Dim2 = c(mfa_auto$eig[2, 3], sum(xc_pct[1:2])),
                         r_Dim1 = c(1, abs(cor(sv$u[, 1], mfa_auto$ind$coord[, 1]))), r_Dim2 = c(1, abs(cor(sv$u[, 2], mfa_auto$ind$coord[, 2]))))
knitr::kable(XCHECK, digits = 3)
```

```{r rev-axes}
# REVISION 2026: empirical interpretation of the axes (primary model)
m <- mfa_auto
# rows of the MFA follow inp_global$meta (groups in alphabetical order); align the item matrix explicitly
row_key <- match(paste(inp_global$meta$respondent_group, inp_global$meta$respondent_id), paste(PARTICIPANTS$respondent_group, PARTICIPANTS$n))
stopifnot(!anyNA(row_key), !anyDuplicated(row_key))
pl_items <- PARTICIPANTS[row_key, item_map$code]
AX_GROUPS <- tibble::tibble(block = rownames(m$group$coord), coord_Dim1 = m$group$coord[, 1], coord_Dim2 = m$group$coord[, 2], contrib_Dim1 = m$group$contrib[, 1],
                            contrib_Dim2 = m$group$contrib[, 2], cos2_Dim1 = m$group$cos2[, 1], cos2_Dim2 = m$group$cos2[, 2])
AX_COMP <- tibble::tibble(component = rownames(m$quanti.var$cor), cor_Dim1 = m$quanti.var$cor[, 1], cor_Dim2 = m$quanti.var$cor[, 2],
                          contrib_Dim1 = m$quanti.var$contrib[, 1], contrib_Dim2 = m$quanti.var$contrib[, 2], cos2_Dim1 = m$quanti.var$cos2[, 1], cos2_Dim2 = m$quanti.var$cos2[, 2])
AX_ITEMS <- item_map %>% dplyr::transmute(code, instrument = ins_lab[instrument], domain = segment, question = label,
               rho_Dim1 = sapply(code, function(v) cor(pl_items[[v]], m$ind$coord[, 1], method = "spearman")),
               rho_Dim2 = sapply(code, function(v) cor(pl_items[[v]], m$ind$coord[, 2], method = "spearman")))
tert <- cut(m$ind$coord[, 1], quantile(m$ind$coord[, 1], c(0, 1/3, 2/3, 1)), include.lowest = TRUE, labels = c("low", "middle", "high"))
AX_ORIENT <- tibble::tibble(tertile_Dim1 = tert, mean_item_score = rowMeans(pl_items)) %>% dplyr::group_by(tertile_Dim1) %>% dplyr::summarise(n = dplyr::n(), mean_item_score = mean(mean_item_score))
eta2 <- function(y, g) { g <- factor(g); 1 - sum(tapply(y, g, function(v) sum((v - mean(v))^2))) / sum((y - mean(y))^2) }
AX_GEOM <- tibble::tibble(factor = c("Respondent group", "Dental school"),
                          eta2_Dim1 = c(eta2(m$ind$coord[, 1], inp_global$meta$respondent_group), eta2(m$ind$coord[, 1], inp_global$meta$ds)),
                          eta2_Dim2 = c(eta2(m$ind$coord[, 2], inp_global$meta$respondent_group), eta2(m$ind$coord[, 2], inp_global$meta$ds)))
AX_BARY <- tibble::tibble(group = inp_global$meta$respondent_group, Dim1 = m$ind$coord[, 1], Dim2 = m$ind$coord[, 2]) %>% dplyr::group_by(group) %>%
  dplyr::summarise(n = dplyr::n(), Dim1_mean = mean(Dim1), Dim1_sd = sd(Dim1), Dim2_mean = mean(Dim2), Dim2_sd = sd(Dim2))
# label tests
AX_TESTS <- tibble::tibble(
  label = c("Dim1 = overall endorsement", "Dim1 = overall endorsement", "Dim1 = overall endorsement", "Dim2 = D-OHR vs CCI contrast", "Dim2 = D-OHR vs CCI contrast", "Dim1 = adoption/maintenance (previous)", "Dim1 = sustainability gradient (previous)"),
  criterion = c("share of the 51 items positively correlated with Dim1", "range of block contributions to Dim1 (%)", "mean item score across Dim1 tertiles (low/middle/high)",
                "mean rho_Dim2 of D-OHR items minus CCI items (sign arbitrary)", "share of items whose rho_Dim2 sign matches their instrument (D-OHR vs CCI)",
                "adoption + maintenance share of Dim1 inertia (%)", "evidence that maintenance/adoption items order participants differently from the other domains on Dim1"),
  value = c(sprintf("%.0f%%", 100 * mean(AX_ITEMS$rho_Dim1 > 0)), sprintf("%.1f-%.1f", min(AX_GROUPS$contrib_Dim1), max(AX_GROUPS$contrib_Dim1)),
            paste(sprintf("%.2f", AX_ORIENT$mean_item_score), collapse = " / "),
            sprintf("%.2f", mean(AX_ITEMS$rho_Dim2[AX_ITEMS$instrument == "D-OHR"]) - mean(AX_ITEMS$rho_Dim2[AX_ITEMS$instrument == "CCI"])),
            { s <- sign(mean(AX_ITEMS$rho_Dim2[AX_ITEMS$instrument == "D-OHR"])); sprintf("%.0f%%", 100 * mean(c(sign(AX_ITEMS$rho_Dim2[AX_ITEMS$instrument == "D-OHR"]) == s, sign(AX_ITEMS$rho_Dim2[AX_ITEMS$instrument == "CCI"]) == -s))) },
            sprintf("%.1f", sum(AX_GROUPS$contrib_Dim1[AX_GROUPS$block %in% c("Adoption", "Maintenance")])),
            sprintf("block contributions are uniform (%.1f-%.1f%%); no domain-specific gradient", min(AX_GROUPS$contrib_Dim1), max(AX_GROUPS$contrib_Dim1))))
knitr::kable(AX_GROUPS, digits = 3); knitr::kable(AX_COMP, digits = 3); knitr::kable(AX_TESTS); knitr::kable(AX_GEOM, digits = 3); knitr::kable(AX_BARY, digits = 2)
```

```{r rev-export}
# REVISION 2026: exports used by the manuscript and response Rmds (no hand-typed numbers downstream)
fig_out <- file.path(proj_dir, "outputs", "figures")
table_out <- file.path(proj_dir, "outputs", "tables")
session_out <- file.path(proj_dir, "session")

dir.create(fig_out, showWarnings = FALSE, recursive = TRUE)
dir.create(table_out, showWarnings = FALSE, recursive = TRUE)
dir.create(session_out, showWarnings = FALSE, recursive = TRUE)
FIG_OUT <- c(
  pdf = file.path(fig_out, "Figure1_final.pdf"),
  png = file.path(fig_out, "Figure1_final.png")
)
ggplot2::ggsave(FIG_OUT[["pdf"]], plot = panel, device = cairo_pdf, width = 11.69, height = 8.27)
ggplot2::ggsave(FIG_OUT[["png"]], plot = panel, width = 11.69, height = 8.27, dpi = 600, bg = "white")
wbx <- function(lst, f) writexl::write_xlsx(lapply(lst, as.data.frame), f)   # REVISION 2026: clean OOXML writer
wbx(list(BLOCK_RETENTION = BLOCKS), file.path(table_out, "MFA_BLOCK_RETENTION.xlsx"))
wbx(list(BLOCK_GROUPS = AX_GROUPS, COMPONENTS = AX_COMP, ITEMS = AX_ITEMS %>% dplyr::arrange(dplyr::desc(abs(rho_Dim1))), LABEL_TESTS = AX_TESTS,
         ORIENTATION = AX_ORIENT, GEOMETRY_ETA2 = AX_GEOM, BARYCENTRES = AX_BARY, EIGENVALUES = as.data.frame(mfa_auto$eig) %>% tibble::rownames_to_column("dim")),
    file.path(table_out, "MFA_AXIS_INTERPRETATION.xlsx"))
wbx(list(SENSITIVITY = SENS, BOOTSTRAP = BOOT, CROSS_CHECK = XCHECK), file.path(table_out, "MFA_SENSITIVITY.xlsx"))
wbx(list(TABLE1_BY_DS = T1_ds, TABLE1_TOTALS = T1_tot, TABLE1_PER_DS = T1_perds, TABLE2 = T2, TABLE3_POOLED = T3_pooled, ITEMS = ITEMS,
         ADHERENCE_FLOW = ADH_FLOW, ADHERENCE_CRITERIA = ADH_CRIT, CCI_VS_DOHR = INS_MEANS), file.path(table_out, "CORRECTED_TABLES_DATA.xlsx"))
saveRDS(list(T1_ds = T1_ds, T1_tot = T1_tot, T1_perds = T1_perds, T2 = T2, ITEMS = ITEMS, T3_pooled = T3_pooled, INS_MEANS = INS_MEANS,
             ADH = ADH, ADH_FLOW = ADH_FLOW, ADH_CRIT = ADH_CRIT, ADH_TOT = ADH_TOT, BLOCKS = BLOCKS, SENS = SENS, BOOT = BOOT, XCHECK = XCHECK,
             eig = mfa_auto$eig, eig_stu = mfa_stu$eig, eig_fixed = mfa_fixed$eig, eig_classic = mfa_classic$eig,
             AX_GROUPS = AX_GROUPS, AX_COMP = AX_COMP, AX_ITEMS = AX_ITEMS, AX_TESTS = AX_TESTS, AX_ORIENT = AX_ORIENT, AX_GEOM = AX_GEOM, AX_BARY = AX_BARY,
             n_zero_var = n_zero_var, n_students = sum(PARTICIPANTS$respondent_group == "Students"), n_educators = sum(PARTICIPANTS$respondent_group == "Educators"),
             item_map = item_map, alpha_segment = alpha_segment, boot_B = B, seed = 20260929),
        file.path(restricted_out, "analysis_results.rds"))
writeLines(
  c(
    paste("Rendered:", format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")),
    "Seed: 20260929",
    capture.output(sessionInfo())
  ),
  file.path(session_out, "sessionInfo.txt")
)

list.files(file.path(proj_dir, "outputs"), recursive = TRUE)
```

```{r rev-session}
sessionInfo()
```
